### 2020-06-10 manuscript/manuscript.Rmd
most of manuscript laid out
still need appendix
clean up table 1

---
title: Multi-scale assessment of a grassland productivity model
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'Biogeosciences')`"
author:
  - given_name: Shawn D.
    surname: Taylor
    affiliation: 1
    email: shawn.taylor@usda.gov
    corresponding: true
  - given_name: Dawn M.
    surname: Browning
    affiliation: 1
    email: dawn.browning@usda.gov
# If you have more than one corresponding author, add them manually using the following structure:
# Two authors: \correspondence{Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)}
# Three authors or more: \correspondence{Daniel Nüst (daniel.nuest@uni-muenster.de), Josiah Carberry j.carberry@orcid.org), and Markus Konkol (m.konkol@wwu.de)}
# If the following line is uncommented, the "corresponding: true" above are ignored
#correspongdingauthors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
affiliation:
  - code: 1
    address: U.S. Department of Agriculture, Agricultural Research Service, Jornada Experimental Range, New Mexico State University, Las Cruces, New Mexico, United States

abstract: |
  Grasslands provide a large amount of ecosystem services globally and forecasting grassland productivity in the coming decades will provide valuable information to land managers. Productivity models can be well-calibrated at local scales, but generally have some maximum spatial extent in which they perform well. Here we evaluate a grassland productivity model to find the optimal spatial extent for parameterization, and also for subsequently applying it in future forecasts for North America. We also evaluated the model on new vegetation types to ascertain its potential generality. We find the model most suitable when incorporating only grasslands, as opposed to also including agriculture and shrublands, and only in the Great Plains and Eastern Temperate Forest ecoregions of North America. The model was not well suited to grasslands in North American Deserts or Northwest Forest ecoregions. It also performed poorly in agriculture vegetation, likely due management activities, and shrubland vegetation, likely because the model lacks representation of deep water pools. This work allows us to perform long-term forecasts in areas where model performance has been verified, with gaps filled in by future modelling efforts.

bibliography: refs
running:
  title: Extended PhenoGrass Evaluation
  author: Taylor & Browning
# This section is mandatory even if you declare that no competing interests are present.
competinginterests: |
  The authors declare no competing interests.
# OPTIONAL:
algorithms: true
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed.
copyrightstatement: |
  Copyright 2020 The Authors
### The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
### It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
availability:
  codedata: |
    All code and data used in the analysis is available in the repository at https://github.com/sdtaylor/PhenograssReplication, the PhenoGrass model is implemented in a python package https://github.com/sdtaylor/GrasslandModels. Both are archived permanently on Zenodo (link).
  sample: |
    use this section when having geoscientific samples available
authorcontribution: |
  Shawn concieved of the project. Shawn and Dawn wrote the paper.
disclaimer: |
  None
acknowledgements: |
  None
appendix: |
 TODO
output:
  rticles::copernicus_article: default
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
---

\introduction

Grassland systems span nearly 30% of the global land surface [@adams1990] and play a prominent role in terrestrial carbon cycles [@parton2012]. Grasslands in North America provide a large proportion of food and fiber agricultural products for the region. Annual productivity of grasslands in central and western North America is driven in large part by precipitation [@sala2012]. Future changes in the amount, intensity, and timing of precipitation will be heterogeneous across North America [@easterling2017], resulting in heterogeneous changes to grassland productivity. For example, even with consistent shifts in climate, different locations can experience different changes in productivity due to local-scale responses [@zhang2011;@sala2012;@knapp2017]. This highlights the need for models which can be resolved at small spatial and temporal scales, thus making long-term grassland productivity forecasts as informative as possible.

There are several potential limitations in the underlying productivity models which can drive such a forecast. Process-based models parameterized with observed data have limited transferability beyond the spatial extent from which their training data came [taylor2019]. For any location the most accurate model will be one which was parameterized from locally collected data, yet these site-specific models will not generalize to new locations [@basler2016]. Incorporating more, and diverse, locations into the model building process will allow it to be more generalizable, yet this comes at a cost of decreased proficiency at all locations [@garcia-mozo2008; @basler2016]. Thus there is an optimal extent in the building and subsequent application of productivity models, which depends on a tradeoff between proficiency at the local scale and applicability at the larger scale.

Here we evaluate a productivity model with the intention of it driving long-term forecasts. The PhenoGrass model developed by @hufkens2016 is a pulse-response productivity model with temperature and precipitation as the primary drivers. The model is parameterized using observations from the Phenocam network, which have a small resolution  (footprints of < 1ha), sub-daily sampling and sites across all major biomes. These attributes make the PhenoGrass model potentially widely applicable. We expand on the evaluation of the original study by using 445 additional site-years of data from 77 new sites. We test the model’s performance across varying combinations of North American ecoregions and vegetation types to find an optimal spatial extent and/or ecoregion for which to parameterize and apply the model. Finally we address where the model performs poorly and how productivity forecasts for these areas could be implemented or improved.

# Methods

## PhenoGrass Model

The PhenoGrass model is an ecohydrology model which has interacting state variables for soil water, plant available water, and plant fractional cover [@hufkens2016]. Model inputs are daily precipitation, temperature, potential evapotranspiration (derived from the Hargreaves equation [@hargreaves1985]), and solar radiation. The primary output is fractional vegetation cover. The original model form, derived in Choler et 2010 and Choler et al. 2011, used only temperature and potential evapotranspiration and was parameterized using satellite-derived NDVI data. @hufkens2016 expanded on the original Choler model by incorporating growth and sensece restraints from temperature and solar radiation, and also included a scaling factor to convert Phenocam GCC data to a fractional cover estimate. @hufkens2016 evaluated the PhenoGrass model using 14 grassland Phenocam sites across Western North America with a total of 34 site years. They found the modelled fractional cover correlated well with annual productivity at both a daily and annual timescale.

## Phenocam Data

The Phenocam network is a global network of fixed, near-surface cameras capturing true-color images of vegetation throughout the day [@richardson2018a]. Using a ratio of the three RGB bands a greenness metric (GCC) is calculated from each image, resulting in a daily scale time series of canopy greeness. GCC is a unitless metric which is highly correlated with satellite derived NDVI [@richardson2018b] and flux tower derived primary productivity [@yan2019; @toomey2015]. Each Phenocam image is subset to one to several different plant vegetation types based on the field of view. These regions of interest (ROI) serve as the basis for the GCC calculation and subsequent post-processing [@seyednasrollah2019].

We downloaded all Phenocam data with ROIs of the grasslands (GR), shrublands (SH), and agricultural (AG) vegetation types for the years 2012 to 2018, totalling 91 sites and 479 site-years (Fig. 1). As input to the PhenoGrass model we used the 3-day smoothed GCC scaled, for each ROI, from 0-1. In the model parameterization each ROI time series is further transformed to a fractional cover estimate using the local mean annual precipitation (MAP) combined with a scaling factor [@hufkens2016; @donohue2013].

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Locations of Phenocam sites. Color indicates the vegetation type represented at each site. Vegetation type is defined by the PhenoCam Network. Shading indicates E.P.A. North American Level 1 Ecoregions."}
knitr::include_graphics("figs/fig1_map.png")
```

## Environmental Data

For historic precipitation and temperature we used the daily 4-km resolution Daymet dataset [@daymet]. Climate time series were extracted for the pixel at the location of each phenocam tower. Daily mean temperature was calculated as the average between the Daymet daily minimum and maximum temperature, and smoothed with a 15 day moving average. Potential evapotranspiration was calculated using the Hargreaves equation [@hargreaves1985]. Soil wilting point and field capacity were extracted at each Phenocam location from a global dataset [@soil2000].

## Model Evaluation

We used a three-step process to evaluate the PhenoGrass model at different spatial extents. First, we parameterized a suite of 15 PhenoGrass models comprising four different extents to find the most appropriate one to make forecasts (Fig. 1). The largest extent used all Phenocam locations described above (84 sites). Next were all sites, respectively, within the three vegetation types (grasslands, shrublands, and agricultural), and all sites, respectively, within each of the four ecoregions (North American Deserts, Great Plains, NW Forests, Eastern Temperate Forests) [@omernik2014]. Finally, we parameterized models for each vegetation type within each ecoregion (eg. All grassland locations within the Great Plains ecoregion). All sets of parameterized models were limited to have at least five sites.

Second, using the same combinations of sites described above, we applied each model parameterization to its respective extent as well as each one below it. For example, sites within the grassland vegetation type were tested using the grassland only model as well as the all site  model. All sites within the Great Plains ecoregion were tested using its own model as well as the all site model. The shrubland vegetation sites within the N.A. Deserts ecoregion were tested using its respective model, the N.A. Deserts, and the all site model. There was no cross-validation using out of sample data in this step as it would have been computationally expensive. Rather, these fitting data metrics were treated as a best case scenario in what each model parameterization can achieve. From these results we used a threshold to select which model to evaluate further using cross-validation at the appropriate scale. For the error metric we chose the coefficient of determination calculated for each site, and averaged across all sites in the respective evaluation. We used a threshold of 0.65, which is viewed as "acceptable" for time-series models [@ritter2013].

```{r, out.width = "8.3cm", echo = FALSE, fig.cap = "Scaling representation of the 15 model parameterizations. Numbers in parentheses represent the number of sites and site years, respectively. Each model uses a different subset of sites ranging from the entire dataset (All Site Model) to one vegetation type within an ecoregion (e.g., Eastern Temperate Forest Grasslands)."}
knitr::include_graphics("figs/fig2_scales.png")
```

Third, models which exceed the threshold described above were subject to further evaluation. For each model we performed a leave one out cross-validation, where the model was re-fit with one Phenocam site not included in the training data, and then evaluated against the left out site. In this step a scaling coefficient to link mean annual precipitation with PhenoCam GCC was held constant at the value obtained in step two. The resulting coefficient of determination is the average among all modelled sites using their respective out of sample test.

All phenocam data was downloaded using the phenocamr R package [@hufkens2018]. Other packages used in the R 3.6 language were dplyr [@dplyr], tidyr [@tidyr], ggplot2 [@ggplot2], daymetr [@hufkens2018], rgdal [@rgdal2019], and sf [@sf2018]. Python 3.7 packages included scipy [@scipy2020], numpy [@numpy2011], pandas [@pandas], and dask [@dask].  All code and data used in the analysis is available in the repository at https://github.com/sdtaylor/PhenograssReplication, the PhenoGrass model is implemented in a python package https://github.com/sdtaylor/GrasslandModels. Both are archived permanently on Zenodo (link).

# Results

At the largest extent, where the phenograss model was parameterized using all 91 sites, the model performed poorly with an average R\textsuperscript{2} value among sites of 0.33. The error of this largest extent model when applied across different subsets of sites (Table 1, leftmost column) shows that agricultural and North American desert sites contribute most to this low R\textsuperscript{2}, while some grassland sites performed well in this regard. Models built at the ecoregion or vegetation type extent also performed poorly overall. The best model performance was achieved when models were built using a single vegetation type subset to ecoregions.

The grassland vegetation type, subset to specific ecoregions, predominantly outperformed other iterations of the PhenoGrass model (Table 1). Models built using grasslands within the Eastern Temperate Forest and Great Plains ecoregions had the highest average R\textsuperscript{2} values of 0.82 and  0.70, respectively. Using leave one out cross-validation on these two grassland model iterations resulted in similar errors of 0.78 and 0.67 for the Eastern Temperate Forest and Great Plains, respectively. Grasslands in the North American deserts were not modelled well at any extent and had the lowest R\textsuperscript{2} values in the entire analysis.

Agriculture and shrubland sites were poorly modelled at all extents. The highest average R\textsuperscript{2} at any extent was 0.51, and 0.39 for shrubland and agriculture, respectively. Shrublands in North American Deserts improved as the model became more and more localized, but never exceeded the 0.65 threshold.

\begin{table}
\caption{Average site-level coefficient of determination (R\textsuperscript{2}) for each vegetation type and ecoregion group parameterized model applied to its respective extent as well as appropriate smaller extents. Bold indicates when a R\textsuperscript{2} was greater than the “acceptable” threshold of 0.65. Values in parentheses represent the average R\textsuperscript{2} in leave-1-out cross validation.}
\begin{tabular}{l|r|r|r|r}
\hline
 & All Site Model & Ecoregion Model & Vegetation Model & Ecoregion+Vegetation Model\\
\hline
All Sites & 0.33 &  &  & \\
\hline
\textbf{E. Temperate Forests} & 0.43 & 0.48 &  & \\
\hline
\hspace{1em}Agriculture & 0.34 & 0.39 & 0.38 & 0.33\\
\hline
\hspace{1em}Grasslands & 0.73 & 0.79 & 0.65 & 0.82\\
\hline
\textbf{Great Plains} & 0.45 & 0.45 &  & \\
\hline
\hspace{1em}Agriculture & 0.22 & 0.15 & 0.17 & 0.18\\
\hline
\hspace{1em}Grasslands & 0.64 & 0.70 & 0.67 & 0.70\\
\hline
\textbf{N. American Deserts} & -0.11 & 0.14 &  & \\
\hline
\hspace{1em}Grasslands & -0.51 & -0.22 & -0.51 & -0.08\\
\hline
\hspace{1em}Shrublands & 0.15 & 0.37 & 0.40 & 0.51\\
\hline
\textbf{N.W. Forests} & 0.25 & 0.58 &  & \\
\hline
\hspace{1em}Grasslands & 0.30 & 0.61 & 0.36 & 0.64\\
\hline
\textbf{All Agriculture} & 0.27 &  & 0.26 & \\
\hline
\textbf{All Grasslands} & 0.43 &  & 0.45 & \\
\hline
\textbf{All Shrublands} & 0.23 &  & 0.39 & \\
\hline
\end{tabular}
\end{table}

# Discussion

We performed an extensive evaluation of the PhenoGrass model across ecoregions and vegetation types to determine the best extent at which to parameterize and apply the model. We found the model most suitable to grassland vegetation when constrained to the ecoregion level, though it did not perform well in grasslands in the North American desert ecoregion. Shrublands and agriculture were not well represented by the model regardless of the extent. Results from this study will facilitate long-term forecasts of grassland productivity constrained to an appropriate vegetation type and extent.

The PhenoGrass model performed best in grassland sites embedded within ecoregions. Studies using earlier forms of the model applied it exclusively to grasslands [@choler2010, @choler2011, @hufkens2016], and results here confirm that it performs well in grassland vegetation with two exceptions. The model did not work in the desert grasslands, nor did it generalize well when built using all North American grasslands simultaneously. Grasslands in the N.A. Desert biome coexist with shrubs, resulting in complex water usage dynamics described in more detail below. The pulse-response design of PhenoGrass, which makes it well suited in areas with high cover of perennial grass, is likely not applicable when grasses are interspersed with woody plants.

Shrublands were not well modelled at any extent. Dryland shrubs, representing 8 of the 15 shrubland Phenocams analysed here, coexist with grasses by accessing different pools of soil water [@weltzin2000; @muldavin2008], thus have different responses to precipitation and resulting greenness patterns [@browning2017; @yan2019]. A prior form of the PhenoGrass model was designed to work with dryland shrubs [@ogle2004], yet here PhenoGrass showed little effectiveness. The current model form has only a single soil water pool with fluxes from precipitation and evapotranspiration, thus is not well-suited for representing the deeper water pools which shrubs use. Potential improvements would likely need to incorporate a deep soil water pool, in addition to the shallow, which are each utilized by the respective plant functional groups. This has already been implemented in highly parameterized ecohydrology models [cites] and could potentially be used here to make a more generalized PhenoShrub model to apply across large scales. This approach could also help in modelling N.A. Desert grasslands which coexist among shrubs.

Agriculture areas performed poorly with the PhenoGrass model. Management practices of crops artificially increase productivity beyond what would naturally occur, and planting and harvest result in abrupt changes in greenness metrics [@begue2018]. While the results here are unsurprising, to our knowledge this is the first attempt to use near-surface images to drive an agricultural productivity model. We have shown that the PhenoGrass model, designed for natural systems, does not generalize to managed agriculture crops. Future work in using Phenocam data to model agricultural productivity would likely need to incorporate crop specific parameters and management activity, which other cropland modelling systems use [@fritz2019]. The integration of the Phenocam network within the Long-Term Agricultural Research (LTAR) will likely be beneficial for this, as the timing and intensity of management activities or experimental treatments can be incorporated into modelling efforts.

@hufkens2016 originally evaluated the Phenograss model using 14 grassland sites distributed among seven North American ecoregions. In their evaluation they had an average R\textsuperscript{2} of 0.71, while here the model performed poorly when using more than 1 ecoregion. It’s likely that the original 14 grassland sites were ideal locations for the PhenoGrass model, since on average they have a single greenup season every year in the spring or summer (Fig. 3A). The additional 24 grassland sites used in the current study have high seasonal variability and elongated growing seasons (Fig. 3B), and were thus more difficult to represent in a single continental scale grassland model.

```{r, out.width = "12cm", echo = FALSE, fig.cap = "Smoothed time series for all 14 grassland sites used in Hufkens et al 2016 (A) and 24 additional grassland sites added in the current study (B). Each line represents the long-term average Green Chromatic Coordinate (GCC) of a single site across all available years, smoothed using a GAM model."}
knitr::include_graphics("figs/fig3_discussion_avg_site_gcc.png")
```

\conclusions

Replication is an important step in the scientific process, especially given newly available data. Here we have validated prior modelling work and highlighted its limitations. Newer small scale vegetation models can be validated in the same framework and applied to areas where PhenoGrass performs poorly. This can result in a spatial ensemble where the output for any one location and vegetation type is represented by the most appropriate model. Our current work will allow for long-term small scale forecasts of grassland productivity for a large fraction of North America.

### 2020-06-10 make_error_tables.R
bold acceptable values

@@ -32,6 +32,10 @@ error_table = primary_errors %>%
   left_join(error_table_order, by=c('prediction_set'='model_fitting_sets')) %>%
   arrange(row_order) %>%
   mutate(row_label = cell_spec(row_label, 'latex', bold=(row_order %in% cell_type_1))) %>%
+   mutate(allsites = cell_spec(allsites, 'latex', bold=(allsites>=0.65))) %>%
+   mutate(ecoregion = cell_spec(ecoregion, 'latex', bold=(ecoregion>=0.65))) %>%
+   mutate(`ecoregion-vegtype` = cell_spec(`ecoregion-vegtype`, 'latex', bold=(`ecoregion-vegtype`>=0.65))) %>%
+   mutate(vegtype = cell_spec(vegtype, 'latex', bold=(vegtype>=0.65))) %>%
   select(row_label, allsites, ecoregion, vegtype,  `ecoregion-vegtype`)
   
 # TODO: need cv errors in here for the 2 grassland models

### 2020-06-10 manuscript/manuscript.Rmd
some manuscript updates
fix a few references, bold some table values, take URL out of refs except where there is no DOI to link

@@
 Both are archived permanently on Zenodo (link).
-  sample: |
-    use this section when having geoscientific samples available
 authorcontribution: |
   Shawn concieved of the project. Shawn and Dawn wrote the paper.
-disclaimer: |
-  None
 acknowledgements: |
@@
 appendix: |
- TODO
+  Appendix: TODO
 output:
@@
 There are several potential limitations in the underlying productivity models which can drive such a forecast.
-Process-based models parameterized with observed data have limited transferability beyond the spatial extent from which their training data came [taylor2019].
+Process-based models parameterized with observed data have limited transferability beyond the spatial extent from which their training data came [@taylor2019a].
 For any location the most accurate model will be one which was parameterized from locally collected data, yet these site-specific models will not generalize to new locations [@basler2016].
@@
 Values in parentheses represent the average R\textsuperscript{2} in leave-1-out cross validation.}
-\begin{tabular}{l|r|r|r|r}
+\begin{tabular}{l|l|l|l|l}
 \hline
@@
 \hline
-\hspace{1em}Grasslands & 0.73 & 0.79 & 0.65 & 0.82\\
+\hspace{1em}Grasslands & \textbf{0.73} & \textbf{0.79} & \textbf{0.65} & \textbf{0.82 (0.78)}\\
 \hline
@@
 \hline
-\hspace{1em}Grasslands & 0.64 & 0.70 & 0.67 & 0.70\\
+\hspace{1em}Grasslands & 0.64 & \textbf{0.70} & \textbf{0.67} & \textbf{0.70 (0.67)}\\
 \hline
@@
 The PhenoGrass model performed best in grassland sites embedded within ecoregions.
-Studies using earlier forms of the model applied it exclusively to grasslands [@choler2010, @choler2011, @hufkens2016], and results here confirm that it performs well in grassland vegetation with two exceptions.
+Studies using earlier forms of the model applied it exclusively to grasslands [@choler2010; @choler2011; @hufkens2016], and results here confirm that it performs well in grassland vegetation with two exceptions.
 The model did not work in the desert grasslands, nor did it generalize well when built using all North American grasslands simultaneously.
@@
 Potential improvements would likely need to incorporate a deep soil water pool, in addition to the shallow, which are each utilized by the respective plant functional groups.
-This has already been implemented in highly parameterized ecohydrology models [cites] and could potentially be used here to make a more generalized PhenoShrub model to apply across large scales.
+This has already been implemented in highly parameterized ecohydrology models [@scanlon2005;@lauenroth2014] and could potentially be used here to make a more generalized PhenoShrub model to apply across large scales.
 This approach could also help in modelling N.A.

### 2020-06-11 discussion_figure_site_gcc_comparison.R
spelling error

@@ -10,7 +10,7 @@ hufkin_sites = read_csv('hufkins2016_sites.csv')
 
 site_status = phenocam_sites %>%
   mutate(site_grouping = case_when(
-    phenocam_name %in% hufkin_sites$phenocam_name ~ 'A. Hufkins et al. 2016 Sites',
+    phenocam_name %in% hufkin_sites$phenocam_name ~ 'A. Hufkens et al. 2016 Sites',
     TRUE ~ 'B. Additional Grassland Sites in Current Study')) %>% 
   select(phenocam_name, site_grouping, ecoregion)
 

### 2020-06-11 discussion_figure_site_gcc_comparison.R
size titles to fit

@@ -29,7 +29,7 @@ discusion_figure = ggplot(grassland_gcc, aes(x=doy, y=gcc, color=as.factor(times
   theme_bw(25) +
   theme(legend.position = 'none',
         strip.background = element_blank(),
-        strip.text = element_text(hjust = 0)) +
+        strip.text = element_text(hjust = 0,size=16)) +
   labs(x='Day of Year', y='Avg. GCC')
 
 ggsave('manuscript/figs/fig3_discussion_avg_site_gcc.png', plot=discusion_figure, height = 15, width=30, units = 'cm', dpi=200)  

### 2020-06-15 make_error_tables.R
slight adjustment

@@ -53,7 +53,8 @@ site_table_column_names = c('Ecoregion','Veg Type','Site','All Site Model','Ecor
 
 site_error_table = site_errors %>%
   select(ecoregion_desc, veg_type = roi_type, phenocam_name, allsites, ecoregion, `ecoregion-vegtype`, vegtype) %>%
-  sample_n(20)
+   arrange(ecoregion_desc, veg_type)
 
 kable(site_error_table, 'latex', col.names = site_table_column_names, escape = F) %>%
-  column_spec(1, width = '40em')
+  column_spec(1, width = '10em') %>%
+   column_spec(6, width='10em')

### 2020-06-15 analysis/site_table.R
site appendix table code

@@ -1,0 +1,22 @@
+library(tidyverse)
+library(kableExtra)
+
+ecoregion_info = read_csv('model_fitting_set_info/ecoregion_codes.csv') %>%
+  mutate(ecoregion_desc = str_wrap(ecoregion_desc,10))
+hufkin2016_sites = read_csv('hufkins2016_sites.csv')
+
+site_list = read_csv('site_list.csv') %>%
+  filter(has_processed_data) %>%
+  mutate(phenocam_name = str_remove(phenocam_name, 'DP1.00033')) %>% # Shorten those long neon names
+  mutate(phenocam_name = str_remove(phenocam_name, 'DP1.20002')) %>%
+  mutate(lat = round(lat,2), lon=round(lon,2)) %>%
+  rename(name = phenocam_name) %>%
+  left_join(ecoregion_info, by='ecoregion') %>%
+  select(name,lat,lon,vegetation=roi_type,roi_id,first_date,last_date,site_years,ecoregion = ecoregion_abbr)
+
+site_list = site_list %>%
+  mutate(name = case_when(
+    name %in% hufkin2016_sites$phenocam_name ~ paste0(name,'*'),
+    TRUE ~name))
+
+kable(site_list, 'latex')

### 2020-06-16 manuscript/manuscript.Rmd
add supplement site table

@@
 appendix: |
-  Appendix: TODO
+  \section{Tables in Appendix}
+  \appendixtables
+  \begin{table}
+  \caption{Sites used in analysis. A * indicates the site was also used in the original analysis.}
+  \small
+  \begin{tabular}{l|r|r|l|r|l|l|r|l}
+  \hline
+  name & lat & lon & vegetation & roi id & first date & last date & site years & ecoregion\\
+  \hline
+  ahwahnee & 37.75 & -119.58 & GR & 3000 & 2015-07-28 & 2020-03-05 & 4.6 & NWForests\\
+  \hline
+  archboldavir & 27.18 & -81.22 & AG & 1000 & 2016-11-18 & 2020-03-05 & 3.1 & ETempForests\\
+  \hline
+  archboldavirx & 27.17 & -81.22 & AG & 1000 & 2016-05-16 & 2020-03-05 & 3.6 & ETempForests\\
+  \hline
+  archboldpnot & 27.19 & -81.20 & AG & 1000 & 2016-05-13 & 2020-03-05 & 3.6 & ETempForests\\
+  \hline
+  archboldpnotx & 27.18 & -81.20 & AG & 1000 & 2016-05-16 & 2020-03-05 & 3.6 & ETempForests\\
+  \hline
+  arsgacp1 & 31.51 & -83.62 & AG & 1000 & 2016-05-10 & 2020-03-05 & 3.8 & ETempForests\\
+  \hline
+  arsmnswanlake1 & 45.68 & -95.80 & AG & 1000 & 2015-10-02 & 2020-03-05 & 4.4 & GrPlains\\
+  \hline
+  bullshoals & 36.56 & -93.07 & GR & 1000 & 2013-11-19 & 2020-03-05 & 5.8 & ETempForests\\
+  \hline
+  burnssagebrush & 43.47 & -119.69 & SH & 1000 & 2012-10-13 & 2020-03-05 & 7.3 & NADeserts\\
+  \hline
+  butte* & 45.95 & -112.48 & GR & 1000 & 2009-01-11 & 2020-03-05 & 10.7 & NWForests\\
+  \hline
+  cperagm & 40.84 & -104.77 & AG & 1000 & 2016-05-19 & 2020-03-05 & 3.8 & GrPlains\\
+  \hline
+  cperagm & 40.84 & -104.77 & GR & 1000 & 2016-05-19 & 2020-03-05 & 3.8 & GrPlains\\
+  \hline
+  cpertgm & 40.83 & -104.76 & GR & 1000 & 2016-05-04 & 2020-03-05 & 3.8 & GrPlains\\
+  \hline
+  cperuvb & 40.81 & -104.76 & GR & 1000 & 2015-07-16 & 2020-03-05 & 4.7 & GrPlains\\
+  \hline
+  gatesofthemountains & 46.83 & -111.71 & GR & 2000 & 2011-08-12 & 2019-02-01 & 7.1 & NWForests\\
+  \hline
+  glacier & 48.50 & -113.99 & GR & 1000 & 2009-02-07 & 2020-03-05 & 11.0 & NWForests\\
+  \hline
+  goodwater & 39.23 & -92.12 & AG & 1001 & 2015-09-26 & 2020-03-05 & 4.5 & GrPlains\\
+  \hline
+  grandteton & 43.92 & -110.58 & SH & 1000 & 2015-07-28 & 2020-02-04 & 3.5 & NWForests\\
+  \hline
+  harvardfarmnorth & 42.52 & -72.18 & AG & 1000 & 2015-11-07 & 2020-03-05 & 4.3 & NForest\\
+  \hline
+  harvardfarmsouth & 42.52 & -72.18 & AG & 1000 & 2015-11-07 & 2020-03-05 & 4.3 & NForest\\
+  \hline
+  harvardgarden & 42.53 & -72.19 & AG & 1000 & 2016-06-12 & 2020-03-05 & 3.7 & NForest\\
+  \hline
+  hawbeckereddy & 40.66 & -77.85 & AG & 1000 & 2015-09-23 & 2019-05-14 & 3.4 & ETempForests\\
+  \hline
+  humnokericea & 34.59 & -91.75 & AG & 1000 & 2015-06-25 & 2020-03-05 & 4.2 & ETempForests\\
+  \hline
+  humnokericec & 34.59 & -91.75 & AG & 1000 & 2015-06-25 & 2020-03-05 & 4.5 & ETempForests\\
+  \hline
+  ibp* & 32.59 & -106.85 & GR & 1000 & 2014-02-16 & 2020-03-05 & 6.0 & NADeserts\\
+  \hline
+  ibp* & 32.59 & -106.85 & SH & 1001 & 2014-02-16 & 2020-03-02 & 6.0 & NADeserts\\
+  \hline
+  jasperridge* & 37.40 & -122.22 & GR & 1000 & 2012-03-08 & 2017-03-09 & 5.0 & MWCoastForests\\
+  \hline
+  jasperridge* & 37.40 & -122.22 & GR & 2000 & 2017-03-18 & 2020-03-05 & 3.0 & MWCoastForests\\
+  \hline
+  jerbajada & 32.58 & -106.63 & SH & 1000 & 2014-04-20 & 2020-03-05 & 5.9 & NADeserts\\
+  \hline
+  jernort & 32.62 & -106.79 & SH & 2000 & 2016-10-28 & 2020-03-05 & 3.3 & NADeserts\\
+  \hline
+  jersand & 32.52 & -106.80 & SH & 1000 & 2014-02-28 & 2020-03-05 & 6.0 & NADeserts\\
+  \hline
+  kansas* & 39.06 & -95.19 & GR & 1000 & 2012-12-03 & 2019-12-31 & 6.7 & GrPlains\\
+  \hline
+  kaweah & 36.44 & -118.91 & SH & 1000 & 2011-07-13 & 2019-09-20 & 8.2 & MedCA\\
+  \hline
+  kelloggcorn & 42.44 & -85.32 & AG & 1000 & 2014-05-23 & 2019-10-05 & 4.9 & ETempForests\\
+  \hline
+  \end{tabular}
+  \end{table}
+
+  \begin{table}
+  \caption{Sites used continued}
+  \small
+  \begin{tabular}{l|r|r|l|r|l|l|r|l}
+  \hline
+  name & lat & lon & vegetation & roi id & first date & last date & site years & ecoregion\\
+  \hline
+  kelloggcorn2 & 42.40 & -85.38 & AG & 1000 & 2015-07-16 & 2019-04-11 & 3.2 & ETempForests\\
+  \hline
+  kelloggcorn3 & 42.40 & -85.37 & AG & 1000 & 2015-07-16 & 2020-03-05 & 3.7 & ETempForests\\
+  \hline
+  kelloggcornsoy2 & 42.40 & -85.37 & AG & 1000 & 2015-07-16 & 2020-03-05 & 3.7 & ETempForests\\
+  \hline
+  kelloggmiscanthus & 42.40 & -85.38 & AG & 1000 & 2015-07-16 & 2020-03-05 & 3.7 & ETempForests\\
+  \hline
+  kelloggoldfield & 42.40 & -85.37 & AG & 1000 & 2015-07-16 & 2020-03-05 & 4.2 & ETempForests\\
+  \hline
+  kendall* & 31.74 & -109.94 & GR & 1000 & 2012-07-06 & 2020-03-05 & 7.6 & SouthAridHighliands\\
+  \hline
+  kendall* & 31.74 & -109.94 & SH & 1000 & 2012-08-08 & 2019-11-07 & 7.2 & SouthAridHighliands\\
+  \hline
+  konza* & 39.08 & -96.56 & GR & 1000 & 2012-03-17 & 2019-12-19 & 6.2 & GrPlains\\
+  \hline
+  lethbridge* & 49.71 & -112.94 & GR & 1000 & 2011-12-07 & 2020-03-05 & 8.3 & GrPlains\\
+  \hline
+  luckyhills & 31.74 & -110.05 & SH & 2000 & 2015-01-26 & 2018-06-04 & 3.4 & SouthAridHighliands\\
+  \hline
+  mandanh5 & 46.78 & -100.95 & AG & 1000 & 2015-09-17 & 2020-03-05 & 4.5 & GrPlains\\
+  \hline
+  mandani2 & 46.76 & -100.93 & AG & 1000 & 2016-04-22 & 2020-03-05 & 3.7 & GrPlains\\
+  \hline
+  manilacotton & 35.89 & -90.14 & AG & 1000 & 2016-06-21 & 2020-03-05 & 3.6 & ETempForests\\
+  \hline
+  marena* & 36.06 & -97.21 & GR & 1000 & 2012-06-12 & 2018-06-19 & 5.8 & GrPlains\\
+  \hline
+  mead1 & 41.17 & -96.48 & AG & 1000 & 2016-07-12 & 2020-03-05 & 3.7 & GrPlains\\
+  \hline
+  mead2 & 41.16 & -96.47 & AG & 1000 & 2016-07-12 & 2020-03-05 & 3.6 & GrPlains\\
+  \hline
+  mead3 & 41.18 & -96.44 & AG & 1000 & 2016-07-12 & 2020-03-05 & 3.7 & GrPlains\\
+  \hline
+  meadpasture & 41.14 & -96.46 & AG & 1000 & 2016-07-15 & 2020-03-05 & 3.7 & GrPlains\\
+  \hline
+  monture* & 47.02 & -113.13 & GR & 2000 & 2010-11-04 & 2019-02-01 & 8.0 & NWForests\\
+  \hline
+  mtrobson & 53.03 & -119.20 & GR & 1000 & 2015-02-16 & 2020-03-05 & 4.8 & NWForests\\
+  \hline
+  nationalelkrefuge & 43.49 & -110.74 & GR & 1000 & 2015-08-12 & 2020-03-05 & 4.5 & NWForests\\
+  \hline
+  NEON.D03.DSNY. & 28.13 & -81.44 & GR & 1000 & 2016-12-15 & 2020-03-05 & 3.2 & ETempForests\\
+  \hline
+  NEON.D06.KONA. & 39.11 & -96.61 & AG & 1000 & 2016-05-07 & 2020-03-05 & 3.7 & GrPlains\\
+  \hline
+  NEON.D06.KONZ. & 39.10 & -96.56 & GR & 1000 & 2017-02-25 & 2020-03-05 & 3.0 & GrPlains\\
+  \hline
+  NEON.D09.WOOD. & 47.13 & -99.24 & GR & 1000 & 2016-12-18 & 2020-03-05 & 3.2 & GrPlains\\
+  \hline
+  NEON.D10.ARIK. & 39.76 & -102.45 & GR & 1000 & 2016-12-18 & 2020-03-05 & 3.1 & GrPlains\\
+  \hline
+  NEON.D10.CPER. & 40.82 & -104.75 & GR & 1000 & 2016-06-30 & 2020-03-05 & 3.7 & GrPlains\\
+  \hline
+  NEON.D10.STER. & 40.46 & -103.03 & AG & 1000 & 2016-12-18 & 2020-03-05 & 3.2 & GrPlains\\
+  \hline
+  NEON.D11.OAES. & 35.41 & -99.06 & GR & 1000 & 2017-02-28 & 2020-03-05 & 3.0 & GrPlains\\
+  \hline
+  NEON.D13.MOAB. & 38.25 & -109.39 & GR & 1000 & 2017-02-25 & 2020-03-05 & 3.0 & NADeserts\\
+  \hline
+  NEON.D14.JORN. & 32.59 & -106.84 & GR & 1000 & 2017-02-25 & 2020-03-05 & 3.0 & NADeserts\\
+  \hline
+  NEON.D14.SRER. & 31.91 & -110.84 & SH & 1000 & 2017-02-25 & 2020-03-05 & 3.0 & NADeserts\\
+  \hline
+  NEON.D15.ONAQ. & 40.18 & -112.45 & GR & 1000 & 2016-12-18 & 2020-03-05 & 3.2 & NADeserts\\
+  \hline
+  NEON.D15.ONAQ. & 40.18 & -112.45 & SH & 1001 & 2016-12-18 & 2020-03-05 & 3.2 & NADeserts\\
+  \hline
+  oakville & 47.90 & -97.32 & GR & 1000 & 2014-08-06 & 2020-03-05 & 4.6 & GrPlains\\
+  \hline
+  \end{tabular}
+  \end{table}
+
+  \begin{table}
+  \caption{Sites used continued}
+  \small
+  \begin{tabular}{l|r|r|l|r|l|l|r|l}
+  \hline
+  name & lat & lon & vegetation & roi id & first date & last date & site years & ecoregion\\
+  \hline
+  pointreyes & 38.00 & -123.02 & SH & 1000 & 2004-01-23 & 2019-03-18 & 13.3 & MedCA\\
+  \hline
+  rosemountnprs & 44.68 & -93.07 & AG & 1000 & 2015-10-29 & 2020-03-05 & 4.3 & GrPlains\\
+  \hline
+  sevilletagrass & 34.36 & -106.70 & GR & 1000 & 2014-11-07 & 2020-03-05 & 5.1 & NADeserts\\
+  \hline
+  sevilletashrub & 34.33 & -106.74 & SH & 1000 & 2014-10-29 & 2020-03-05 & 5.0 & NADeserts\\
+  \hline
+  silverton & 45.00 & -122.69 & AG & 1000 & 2013-07-22 & 2020-02-10 & 5.9 & NWForests\\
+  \hline
+  smokypurchase & 35.59 & -83.07 & GR & 2000 & 2016-01-02 & 2020-03-05 & 4.0 & ETempForests\\
+  \hline
+  southerngreatplains & 36.61 & -97.49 & AG & 1000 & 2012-05-16 & 2020-03-05 & 7.5 & GrPlains\\
+  \hline
+  spruceT6P16E & 47.51 & -93.45 & SH & 1000 & 2015-08-24 & 2020-03-05 & 4.6 & NForest\\
+  \hline
+  stjones & 39.09 & -75.44 & GR & 1000 & 2015-09-20 & 2020-03-05 & 4.4 & ETempForests\\
+  \hline
+  sweetbriargrass & 37.56 & -79.09 & AG & 1000 & 2016-03-23 & 2020-03-05 & 3.8 & ETempForests\\
+  \hline
+  teddy & 46.89 & -103.38 & GR & 1000 & 2010-01-02 & 2019-03-27 & 8.5 & GrPlains\\
+  \hline
+  teddy & 46.89 & -103.38 & SH & 1000 & 2003-05-11 & 2019-03-27 & 14.4 & GrPlains\\
+  \hline
+  tonzi* & 38.43 & -120.97 & GR & 1000 & 2011-10-26 & 2020-02-13 & 7.3 & MedCA\\
+  \hline
+  turkeypointenf02 & 42.66 & -80.56 & AG & 1000 & 2012-05-01 & 2020-03-05 & 7.5 & ETempForests\\
+  \hline
+  twitchell & 38.11 & -121.65 & AG & 1000 & 2011-11-16 & 2017-04-05 & 4.0 & MedCA\\
+  \hline
+  uiefmaize & 40.06 & -88.20 & AG & 1000 & 2008-11-06 & 2020-03-02 & 11.3 & ETempForests\\
+  \hline
+  uiefmiscanthus & 40.06 & -88.20 & AG & 1000 & 2008-11-12 & 2018-04-29 & 9.2 & ETempForests\\
+  \hline
+  uiefprairie* & 40.06 & -88.20 & GR & 1000 & 2008-10-22 & 2020-03-02 & 11.3 & ETempForests\\
+  \hline
+  uiefswitchgrass & 40.06 & -88.20 & AG & 1000 & 2008-10-22 & 2020-03-02 & 10.6 & ETempForests\\
+  \hline
+  usgseros & 43.73 & -96.62 & GR & 1000 & 2014-09-11 & 2017-10-02 & 3.1 & GrPlains\\
+  \hline
+  uwmfieldsta & 43.39 & -88.02 & GR & 1000 & 2013-03-15 & 2020-03-05 & 7.0 & ETempForests\\
+  \hline
+  vaira* & 38.41 & -120.95 & GR & 1000 & 2011-10-17 & 2020-02-28 & 7.7 & MedCA\\
+  \hline
+  \end{tabular}
+  \end{table}
+
 output:

### 2020-06-16 manuscript/manuscript.Rmd
final small updates before submission

@@
 Productivity models can be well-calibrated at local scales, but generally have some maximum spatial extent in which they perform well.
-Here we evaluate a grassland productivity model to find the optimal spatial extent for parameterization, and also for subsequently applying it in future forecasts for North America.
+Here we evaluate a grassland productivity model to find the optimal spatial extent for parameterization, and thus for subsequently applying it in future forecasts for North America.
 We also evaluated the model on new vegetation types to ascertain its potential generality.
@@
 The model was not well suited to grasslands in North American Deserts or Northwest Forest ecoregions.
-It also performed poorly in agriculture vegetation, likely due management activities, and shrubland vegetation, likely because the model lacks representation of deep water pools.
+It also performed poorly in agriculture vegetation, likely due to management activities, and shrubland vegetation, likely because the model lacks representation of deep water pools.
 This work allows us to perform long-term forecasts in areas where model performance has been verified, with gaps filled in by future modelling efforts.
@@
 We expand on the evaluation of the original study by using 445 additional site-years of data from 77 new sites.
-We test the model’s performance across varying combinations of North American ecoregions and vegetation types to find an optimal spatial extent and/or ecoregion for which to parameterize and apply the model.
+We test the model’s performance across varying combinations of North American ecoregions and vegetation types to find an optimal spatial extent in which to parameterize and apply the model.
 Finally we address where the model performs poorly and how productivity forecasts for these areas could be implemented or improved.
@@
 The PhenoGrass model is an ecohydrology model which has interacting state variables for soil water, plant available water, and plant fractional cover [@hufkens2016].
-Model inputs are daily precipitation, temperature, potential evapotranspiration (derived from the Hargreaves equation [@hargreaves1985]), and solar radiation.
+Model inputs are daily precipitation, temperature, potential evapotranspiration (derived from the Hargreaves equation, @hargreaves1985), and solar radiation.
 The primary output is fractional vegetation cover.
@@
 2011, used only temperature and potential evapotranspiration and was parameterized using satellite-derived NDVI data.
-@hufkens2016 expanded on the original Choler model by incorporating growth and sensece restraints from temperature and solar radiation, and also included a scaling factor to convert Phenocam GCC data to a fractional cover estimate.
+@hufkens2016 expanded on the original Choler model by incorporating growth and senescence restraints from temperature and solar radiation, and also included a scaling factor to convert Phenocam G\textsubscript{cc} data to a fractional cover estimate.
 @hufkens2016 evaluated the PhenoGrass model using 14 grassland Phenocam sites across Western North America with a total of 34 site years.
@@
 The Phenocam network is a global network of fixed, near-surface cameras capturing true-color images of vegetation throughout the day [@richardson2018a].
-Using a ratio of the three RGB bands a greenness metric (GCC) is calculated from each image, resulting in a daily scale time series of canopy greeness.
-GCC is a unitless metric which is highly correlated with satellite derived NDVI [@richardson2018b] and flux tower derived primary productivity [@yan2019; @toomey2015].
+Using a ratio of the three RGB bands a greenness metric (green chromatic coordinate, G\textsubscript{cc}) is calculated from each image, resulting in a daily scale time series of canopy greeness.
+G\textsubscript{cc} is a unitless metric which is highly correlated with satellite derived NDVI [@richardson2018b] and flux tower derived primary productivity [@yan2019; @toomey2015].
 Each Phenocam image is subset to one to several different plant vegetation types based on the field of view.
-These regions of interest (ROI) serve as the basis for the GCC calculation and subsequent post-processing [@seyednasrollah2019].
+These regions of interest (ROI) serve as the basis for the G\textsubscript{cc} calculation and subsequent post-processing [@seyednasrollah2019].
 
 We downloaded all Phenocam data with ROIs of the grasslands (GR), shrublands (SH), and agricultural (AG) vegetation types for the years 2012 to 2018, totalling 91 sites and 479 site-years (Fig.
-1).
-As input to the PhenoGrass model we used the 3-day smoothed GCC scaled, for each ROI, from 0-1.
+1, Table A1).
+As input to the PhenoGrass model we used the 3-day smoothed G\textsubscript{cc} scaled, for each ROI, from 0-1.
 In the model parameterization each ROI time series is further transformed to a fractional cover estimate using the local mean annual precipitation (MAP) combined with a scaling factor [@hufkens2016; @donohue2013].
@@
 First, we parameterized a suite of 15 PhenoGrass models comprising four different extents to find the most appropriate one to make forecasts (Fig.
-1).
+2).
 The largest extent used all Phenocam locations described above (84 sites).
@@
 For example, sites within the grassland vegetation type were tested using the grassland only model as well as the all site  model.
-All sites within the Great Plains ecoregion were tested using its own model as well as the all site model.
+All sites within the Great Plains ecoregion were tested using the Great Plains ecoregion model as well as the all site model.
 The shrubland vegetation sites within the N.A.
 Deserts ecoregion were tested using its respective model, the N.A.
-Deserts, and the all site model.
+Deserts ecoregion model, and the all site model.
 There was no cross-validation using out of sample data in this step as it would have been computationally expensive.
-Rather, these fitting data metrics were treated as a best case scenario in what each model parameterization can achieve.
+Rather, error metrics from these in-sample tests were treated as a best case scenario in what each model parameterization can achieve.
 From these results we used a threshold to select which model to evaluate further using cross-validation at the appropriate scale.
-For the error metric we chose the coefficient of determination calculated for each site, and averaged across all sites in the respective evaluation.
+For the error metric we chose the coefficient of determination (R\textsuperscript{2}) calculated for each site, and averaged across all sites in the respective evaluation.
 We used a threshold of 0.65, which is viewed as "acceptable" for time-series models [@ritter2013].
@@
 For each model we performed a leave one out cross-validation, where the model was re-fit with one Phenocam site not included in the training data, and then evaluated against the left out site.
-In this step a scaling coefficient to link mean annual precipitation with PhenoCam GCC was held constant at the value obtained in step two.
+In this step a scaling coefficient to link mean annual precipitation with PhenoCam G\textsubscript{cc} was held constant at the value obtained in step two.
 The resulting coefficient of determination is the average among all modelled sites using their respective out of sample test.
@@
 Models built at the ecoregion or vegetation type extent also performed poorly overall.
-The best model performance was achieved when models were built using a single vegetation type subset to ecoregions.
+The best model performance was achieved when models were built using a single vegetation type subset to a single ecoregion.
 
@@
 ```{r, out.width = "12cm", echo = FALSE, fig.cap = "Smoothed time series for all 14 grassland sites used in Hufkens et al 2016 (A) and 24 additional grassland sites added in the current study (B).
-Each line represents the long-term average Green Chromatic Coordinate (GCC) of a single site across all available years, smoothed using a GAM model."}
+Each line represents the long-term average Green Chromatic Coordinate of a single site across all available years, smoothed using a GAM model."}
 knitr::include_graphics("figs/fig3_discussion_avg_site_gcc.png")

### 2020-08-24 fit_model_cv.py
model CV for recent runs
doing leave 1 out cross validation on grassland models
for NWForests, ETempForests, and  Great Plains

@@ -30,7 +30,7 @@ models_to_validate = {'ETempForests_GR' : "ecoregion-vegtype_ETempForests_GR_Phe
                       'NWForests_GR'    : "ecoregion-vegtype_NWForests_GR_NaiveMAPCorrected_cd499f8473ab478e.json"}
 model_folder = 'fitted_models/2020-07-29_0a82f11673e54a9c/'
 
-model_fitting_note = 'first go at leave 1 out CV with the 4 GR models'
+model_fitting_note = 'CV model runs on the three GR models'
 
 
 ###############################################3
@@ -40,7 +40,7 @@ de_popsize     = 400
 de_maxiter     = 10000
 
 # Dask/ceres stuff
-chunks_per_job = 15
+chunks_per_job = 30
 parameters_count = 7 # phenograss parameters since its the most intensive. 
 
 use_default_workers = False

### 2020-08-26 analysis/error_figures.R
fix fig 3,4 errors
R^2 and rmse should be at the site level and then averaged.

@@ -17,11 +17,9 @@ phenocam_site_info = read_csv('site_list.csv') %>%
 get_label_stats = function(df){
   # This needs to accept a group_by() output
   df %>%
-  summarise(rmse = sqrt(mean( (fCover_observed - fCover_predicted)^2 ,na.rm=T)),
-            r2   = 1  - sum((fCover_observed - fCover_predicted)^2,na.rm=T) / sum((fCover_observed - mean(fCover_observed,na.rm=T))^2,na.rm=T),
-            #mae  = mean(abs(gcc_observed - gcc_predicted),na.rm=T),
-            n=n(),
-            n_timeseries = n_distinct(timeseries_id)) %>%
+    summarise(rmse = mean(rmse),
+              r2 =   mean(r2),
+              n_timeseries = n_distinct(timeseries_id)) %>%
     ungroup() %>%
     mutate(label_text = paste0('R2: ',round(r2,2),'\n','RMSE: ',round(rmse,2),'\n',n_timeseries,' sites'))
 }
@@ -40,6 +38,12 @@ allsite_veg_predictions = all_predictions %>%
 allsite_veg_predictions$fitting_set = factor(allsite_veg_predictions$fitting_set, levels = fitting_levels, labels = nice_labels)
 
 allsite_veg_predictions_error_text = allsite_veg_predictions %>%
+  group_by(fitting_set, timeseries_id) %>%
+  summarise(rmse = sqrt(mean( (fCover_observed - fCover_predicted)^2 ,na.rm=T)),
+            r2   = 1  - sum((fCover_observed - fCover_predicted)^2,na.rm=T) / sum((fCover_observed - mean(fCover_observed,na.rm=T))^2,na.rm=T),
+            n=n(),
+            percent_na=mean(is.na(fCover_observed))) %>%
+  ungroup() %>%
   group_by(fitting_set) %>%
   get_label_stats()
 
@@ -81,10 +85,14 @@ ecoregion_vegtype_predictions = ecoregion_vegtype_predictions %>%
   mutate(facet_label = paste(ecoregion_desc,vegtype,sep=' - '))
 
 ecoregion_vegtype_predictions_error_text = ecoregion_vegtype_predictions %>%
+  group_by(facet_label, timeseries_id) %>%
+  summarise(rmse = sqrt(mean( (fCover_observed - fCover_predicted)^2 ,na.rm=T)),
[…]

### 2020-08-26 apply_models.py
update to newest model set

@@ -14,16 +14,10 @@ phenocam_info = pd.read_csv('site_list.csv')
 phenocam_info = phenocam_info[phenocam_info.has_processed_data]
 ######################################################
 # fitted models to apply
-model_set_id = '4dac8b702c3241eb'
+model_set_id = '0a82f11673e54a9c'
 model_set = load_models.load_model_set(model_set_id)
 
 fitted_models = [{'model':m,'model_name':n,'fitting_years':'20012009'} for m,n in zip(model_set['models'],model_set['model_names'])]
-
-# also some naive models fit later on 
-model_set_id = '660659bec01842ec'
-model_set = load_models.load_model_set(model_set_id)
-
-fitted_models.extend([{'model':m,'model_name':n,'fitting_years':'20012009'} for m,n in zip(model_set['models'],model_set['model_names'])])
 
 ######################################################
 # For each model get the predictions for it's training set

### 2020-08-26 manuscript/manuscript.Rmd
zenodo link
Downsize analysis for resubmission (#9)
* remove ecoregion scale from analysis

This simplifies things as aggregatign all sites within
an ecoregion doesn't add much to the analysis.

* simplify by dropping the cross-scale error comparison.

* Add new error figures of observed/prediction plots

* drop pointreyes

its causing errors somehow. Its only used in the allsite
and vegtype-SH models so i'm not losing much.

* drop jasperridge 2000 ROI

just a smaller ROI than the 1000 one

* new supplement figure

highlighting the really bad performing desert grasslands

* table A1 to append

--- 2020-06-10 manuscript/manuscript.Rmd
@@
     All code and data used in the analysis is available in the repository at https://github.com/sdtaylor/PhenograssReplication, the PhenoGrass model is implemented in a python package https://github.com/sdtaylor/GrasslandModels.
-Both are archived permanently on Zenodo (link).
+Both are archived permanently on Zenodo (https://doi.org/10.5281/zenodo.3897319).
 authorcontribution: |
@@
   None
-appendix: |
-  Appendix: TODO
 output:
@@
 All code and data used in the analysis is available in the repository at https://github.com/sdtaylor/PhenograssReplication, the PhenoGrass model is implemented in a python package https://github.com/sdtaylor/GrasslandModels.
-Both are archived permanently on Zenodo (link).
+Both are archived permanently on Zenodo (https://doi.org/10.5281/zenodo.3897319).
 

### 2020-08-26 analysis/NA_desert_grassland_fig.R

@@ -1,0 +1,39 @@
+library(tidyverse)
+
+#####################################################
+# This produces the supplementary figure S1, which shows the predicted and observed
+# timeseries of the 5 sites used in North American Desert grasslands model. 
+######################################################
+
+all_predictions = read_csv('data/full_model_predictions.csv', 
+                           col_types = cols(W=col_number(),
+                                            gcc_observed = col_number(),
+                                            fCover_observed = col_number(),
+                                            Dt=col_number()))  %>%
+  filter(model=='PhenoGrass') %>%
+  filter(fitting_set %in% c('ecoregion-vegtype_NADeserts_GR'))
+
+
+ecoregion_info = read_csv('model_fitting_set_info/ecoregion_codes.csv')
+phenocam_site_info = read_csv('site_list.csv') %>%
+  select(phenocam_name, timeseries_id, roi_type, ecoregion, MAP = MAP_daymet) %>%
+  left_join(ecoregion_info, by='ecoregion') %>%
+  as_tibble()
+
+all_predictions = all_predictions %>%
+  left_join(phenocam_site_info, by='timeseries_id')
+
+
+desert_grassland_fig = all_predictions %>%
+  filter(date >= '2015-01-01') %>%
+ggplot(aes(x=date)) + 
+  geom_point(aes(y=fCover_predicted), color='#D55E00') +
+  geom_point(aes(y=fCover_observed), color='black') +
+  geom_col(aes(y=precip/100), color='#0072B2') + 
+  facet_wrap(~paste(phenocam_name,roi_type,sep='-')) +
+  theme_bw(15) +
+  theme(axis.text = element_text(color='black')) + 
+  labs(x='Date',y='fCover (unitless) and precipitation (decimeters)')
+
+ggsave('manuscript/figs/figS1_desert_grasslands.png',desert_grassland_fig, height = 20, width=30, units='cm', dpi=100)
+             

### 2020-08-26 analysis/analyze_errors.R

@@ -3,194 +3,31 @@ library(tidyverse)
 all_predictions = read_csv('data/full_model_predictions.csv', 
                            col_types = cols(W=col_number(),
                                             gcc_observed = col_number(),
+                                            fCover_observed = col_number(),
                                             Dt=col_number()))
 
-
-errors = all_predictions %>%
+site_errors = all_predictions %>%
   group_by(fitting_set, model, timeseries_id) %>%
-  summarise(rmse = sqrt(mean( (gcc_observed - gcc_predicted)^2 ,na.rm=T)),
-            r2   = 1  - sum((gcc_observed - gcc_predicted)^2,na.rm=T) / sum((gcc_observed - mean(gcc_observed,na.rm=T))^2,na.rm=T),
-            #mae  = mean(abs(gcc_observed - gcc_predicted),na.rm=T),
+  summarise(rmse = sqrt(mean( (fCover_observed - fCover_predicted)^2 ,na.rm=T)),
+            r2   = 1  - sum((fCover_observed - fCover_predicted)^2,na.rm=T) / sum((fCover_observed - mean(fCover_observed,na.rm=T))^2,na.rm=T),
             n=n(),
-            n_timeseries = n_distinct(timeseries_id),
-            percent_na=mean(is.na(gcc_observed))) %>%
+            percent_na=mean(is.na(fCover_observed))) %>%
+  ungroup() 
+
+primary_errors = site_errors %>%
+  filter(model == 'PhenoGrass') %>% # just the phenograss model in the primary things. naive errors will be in supplement
+  group_by(fitting_set, model) %>%
+  summarise(rmse = round(mean(rmse),2),
+            r2 =   round(mean(r2),2),
+            n_timeseries = n_distinct(timeseries_id)) %>%
   ungroup()
 
-fitting_sets = read_csv('model_fitting_set_info/fitting_sets.csv')
-fitting_set_assignments = read_csv('model_fitting_set_info/fitting_set_assignments.csv')
+# add in site years
+fitting_set_info = read_csv('model_fitting_set_info/fitting_sets.csv') %>%
+  select(model_fitting_sets, site_years = total_site_years)
 
-#######################################
-# Calculate summary errors at the landcover/ecoregion scale for the main result table
[…]

### 2020-08-26 analysis/error_figures.R

@@ -1,0 +1,113 @@
+library(tidyverse)
+
+all_predictions = read_csv('data/full_model_predictions.csv', 
+                           col_types = cols(W=col_number(),
+                                            gcc_observed = col_number(),
+                                            fCover_observed = col_number(),
+                                            Dt=col_number()))  %>%
+  filter(model=='PhenoGrass')
+
+ecoregion_info = read_csv('model_fitting_set_info/ecoregion_codes.csv')
+
+phenocam_site_info = read_csv('site_list.csv') %>%
+  select(phenocam_name, timeseries_id, roi_type, ecoregion, MAP = MAP_daymet) %>%
+  left_join(ecoregion_info, by='ecoregion') %>%
+  as_tibble()
+
+get_label_stats = function(df){
+  # This needs to accept a group_by() output
+  df %>%
+    summarise(rmse = mean(rmse),
+              r2 =   mean(r2),
+              n_timeseries = n_distinct(timeseries_id)) %>%
+    ungroup() %>%
+    mutate(label_text = paste0('R2: ',round(r2,2),'\n','RMSE: ',round(rmse,2),'\n',n_timeseries,' sites'))
+}
+
+#######################################################
+# 1:1 plots of all site model and all vegtype models
+#######################################################
+fitting_levels = c('allsites','vegtype_GR','vegtype_AG','vegtype_SH')
+nice_labels    = c('All Sites', 'All Grasslands','All Agriculture', 'All Shrubland')
+
+allsite_veg_predictions = all_predictions %>%
+  filter(fitting_set %in% c('allsites','vegtype_GR','vegtype_AG','vegtype_SH')) %>%
+  filter(!is.na(fCover_observed)) %>%
+  left_join(phenocam_site_info, by='timeseries_id')
+
+allsite_veg_predictions$fitting_set = factor(allsite_veg_predictions$fitting_set, levels = fitting_levels, labels = nice_labels)
+
+allsite_veg_predictions_error_text = allsite_veg_predictions %>%
+  group_by(fitting_set, timeseries_id) %>%
+  summarise(rmse = sqrt(mean( (fCover_observed - fCover_predicted)^2 ,na.rm=T)),
[…]

### 2020-08-26 analysis/make_error_tables.R

@@ -8,53 +8,113 @@ primary_errors = read_csv('results/primary_error_table.csv')
 error_table_order = tribble(
   ~model_fitting_sets, ~row_label, ~row_order, 
    'allsites', 'All Sites',  1,          
-   'ecoregion_ETempForests', 'E. Temperate Forests', 2,
-   'ecoregion-vegtype_ETempForests_AG', 'Agriculture', 3,
-   'ecoregion-vegtype_ETempForests_GR', 'Grasslands', 4,
+   'vegtype_AG', 'All Agriculture', 2,
+   'vegtype_GR', 'All Grasslands', 3,
+   'vegtype_SH', 'All Shrublands', 4,
+  
+   'ETempForests', 'E. Temperate Forests', 5,
+   'ecoregion-vegtype_ETempForests_AG', 'Agriculture', 6,
+   'ecoregion-vegtype_ETempForests_GR', 'Grasslands', 7,
    
-   'ecoregion_GrPlains', 'Great Plains', 5,
-   'ecoregion-vegtype_GrPlains_AG', 'Agriculture', 6,
-   'ecoregion-vegtype_GrPlains_GR', 'Grasslands', 7,
-   'ecoregion_NADeserts', 'N. American Deserts', 8,
-   'ecoregion-vegtype_NADeserts_GR', 'Grasslands',9,
-   'ecoregion-vegtype_NADeserts_SH', 'Shrublands',10,
+   'GrPlains', 'Great Plains', 8,
+   'ecoregion-vegtype_GrPlains_AG', 'Agriculture', 9,
+   'ecoregion-vegtype_GrPlains_GR', 'Grasslands', 10,
+  
+   'NADeserts', 'N. American Deserts', 11,
+   'ecoregion-vegtype_NADeserts_GR', 'Grasslands',12,
+   'ecoregion-vegtype_NADeserts_SH', 'Shrublands',13,
    
-   'ecoregion_NWForests', 'N.W. Forests', 11,
-   'ecoregion-vegtype_NWForests_GR', 'Grasslands', 12,
-   'vegtype_AG',                      'All Agriculture', 13,
-   'vegtype_GR',                      'All Grasslands', 14,
-   'vegtype_SH',                      'All Shrublands', 15,
+   'NWForests', 'N.W. Forests', 14,
+   'ecoregion-vegtype_NWForests_GR', 'Grasslands', 15,
+
 )
 
-cell_type_1 = c(2,5,8,11,13,14,15)
[…]

### 2020-08-26 apply_model_cv.py

@@ -18,7 +18,7 @@ phenocam_info = pd.read_csv('site_list.csv')
 phenocam_info = phenocam_info[phenocam_info.has_processed_data]
 ######################################################
 # fitted models to apply
-model_set_id = 'c41f4f6a115d4c3a'
+model_set_id = '69f5e0932e7b4970'
 model_set = load_models.load_model_set(model_set_id)
 
 fitted_models = [{'model':m,'model_name':n,'fitting_years':'20012009'} for m,n in zip(model_set['models'],model_set['model_names'])]

### 2020-08-26 apply_models.py

@@ -14,7 +14,7 @@ phenocam_info = pd.read_csv('site_list.csv')
 phenocam_info = phenocam_info[phenocam_info.has_processed_data]
 ######################################################
 # fitted models to apply
-model_set_id = '4dac8b702c3241eb'
+model_set_id = '0a82f11673e54a9c'
 model_set = load_models.load_model_set(model_set_id)
 
 fitted_models = [{'model':m,'model_name':n,'fitting_years':'20012009'} for m,n in zip(model_set['models'],model_set['model_names'])]
@@ -27,6 +27,10 @@ all_data = pd.DataFrame()
 for fit_model in fitted_models:
     
     m = fit_model['model']
+    
+    # forget about original Naive model for now
+    if isinstance(m,GrasslandModels.models.naive.Naive):
+        continue
     
     fitting_set = m.metadata['fitting_set']
     timeseries_used = list(literal_eval(m.metadata['timeseries_used'])) # these are stored as a str, '1,2,3,4,5', for clarity

### 2020-08-26 data_prep/generate_model_fitting_sets.R

@@ -25,16 +25,6 @@ ecoregions_vegtype_sets = phenocam_sites %>%
   mutate(model_fitting_sets = paste('ecoregion-vegtype',ecoregion_abbr,roi_type,sep='_')) %>%
   mutate(model_fitting_scale = 'ecoregion-vegtype')
 
-ecoregions_sets = phenocam_sites %>%
-  group_by(ecoregion) %>%
-  summarise(n_phenocams = n(),
-            total_site_years = sum(site_years)) %>%
-  ungroup() %>%
-  filter(n_phenocams>=min_sites_needed) %>%
-  left_join(ecoregion_codes, by='ecoregion') %>%
-  mutate(model_fitting_sets = paste('ecoregion',ecoregion_abbr,sep='_')) %>%
-  mutate(model_fitting_scale = 'ecoregion')
-
 vegtype_sets = phenocam_sites %>%
   group_by(roi_type) %>%
   summarise(n_phenocams = n(),
@@ -47,10 +37,6 @@ vegtype_sets = phenocam_sites %>%
 ##########################
 # Assign timeseries to each set. A timeseries can be in the 3
 # broad catagories only once each.
-ecoregion_set_assignments = ecoregions_sets %>%
-  select(ecoregion, model_fitting_sets) %>%
-  left_join(phenocam_sites, by=c('ecoregion')) %>%
-  select(model_fitting_sets, timeseries_id)
 
 ecoregions_vegtype_set_assignments = ecoregions_vegtype_sets %>%
   select(ecoregion, roi_type, model_fitting_sets) %>%
@@ -67,19 +53,17 @@ allsite_assignments = phenocam_sites %>%
   mutate(model_fitting_sets = 'allsites') %>%
   select(model_fitting_sets, timeseries_id)
 
-all_assignments = ecoregion_set_assignments %>%
-  bind_rows(ecoregions_vegtype_set_assignments) %>%
+all_assignments = allsite_assignments %>%
   bind_rows(vegtype_assignments) %>%
-  bind_rows(allsite_assignments)
+  bind_rows(ecoregions_vegtype_set_assignments)
 
-# Sanity check, each timeseries should not be represented > 4 times
+# Sanity check, each timeseries should not be represented > 3 times
 x= all_assignments %>% count(timeseries_id) %>% pull(n)
-if(any(x>4)) stop('some phenocam timeseries added more than 4 times')
+if(any(x>3)) stop('some phenocam timeseries added more than 3 times')
 
[…]

### 2020-08-26 fit_model_cv.py

@@ -25,11 +25,12 @@ of the models.
 
 ###############################################
 
-models_to_validate = {'ETempForests_GR' : "ecoregion-vegtype_ETempForests_GR_PhenoGrass_4dac8b702c3241eb.json",
-                      'GrPlains_GR'     : "ecoregion-vegtype_GrPlains_GR_PhenoGrass_4dac8b702c3241eb.json"}
-model_folder = 'fitted_models/2020-05-29_4dac8b702c3241eb/'
+models_to_validate = {'ETempForests_GR' : "ecoregion-vegtype_ETempForests_GR_PhenoGrass_cd499f8473ab478e.json",
+                      'GrPlains_GR'     : "ecoregion-vegtype_GrPlains_GR_PhenoGrass_cd499f8473ab478e.json",
+                      'NWForests_GR'    : "ecoregion-vegtype_NWForests_GR_NaiveMAPCorrected_cd499f8473ab478e.json"}
+model_folder = 'fitted_models/2020-07-29_0a82f11673e54a9c/'
 
-model_fitting_note = 'first go at leave 1 out CV with the 4 GR models'
+model_fitting_note = 'CV model runs on the three GR models'
 
 
 ###############################################3
@@ -39,7 +40,7 @@ de_popsize     = 400
 de_maxiter     = 10000
 
 # Dask/ceres stuff
-chunks_per_job = 15
+chunks_per_job = 30
 parameters_count = 7 # phenograss parameters since its the most intensive. 
 
 use_default_workers = False

### 2020-08-26 analysis/discussion_figure_site_gcc_comparison.R
fig3 -> fig5

@@ -32,7 +32,7 @@ discusion_figure = ggplot(grassland_gcc, aes(x=doy, y=gcc, color=as.factor(times
         strip.text = element_text(hjust = 0,size=16)) +
   labs(x='Day of Year', y='Avg. GCC')
 
-ggsave('manuscript/figs/fig3_discussion_avg_site_gcc.png', plot=discusion_figure, height = 15, width=30, units = 'cm', dpi=200)  
+ggsave('manuscript/figs/fig5_discussion_avg_site_gcc.png', plot=discusion_figure, height = 15, width=30, units = 'cm', dpi=200)  
   
 
   # Also while where here parse out what the original ecoregions were

### 2020-08-26 manuscript/manuscript.Rmd
new manuscript text
revisions for "resubmissin" to biogeosciences.
new figures too

@@
 abstract: |
-  Grasslands provide a large amount of ecosystem services globally and forecasting grassland productivity in the coming decades will provide valuable information to land managers.
+  Grasslands provide many important ecosystem services globally and forecasting grassland productivity in the coming decades will provide valuable information to land managers.
 Productivity models can be well-calibrated at local scales, but generally have some maximum spatial extent in which they perform well.
@@
 acknowledgements: |
-  None
+  This research was a contribution from the Long-Term Agroecosystem Research (LTAR) network.
+LTAR is supported by the United States Department of Agriculture.
+The authors acknowledge the USDA Agricultural Research Service (ARS) Big Data Initiative and SCINet high performance computing resources (https://scinet.usda.gov) made available for conducting the research reported in this paper.
+
+  We thank our many collaborators, including site PIs and technicians, for their efforts in support of PhenoCam.
+The development of PhenoCam has been funded by the Northeastern States Research Cooperative, NSF’s Macrosystems Biology program (awards EF-1065029 and EF-1702697), and DOE’s Regional and Global Climate Modeling program (award DE-SC0016011).
+
 output:
@@
 The PhenoGrass model developed by @hufkens2016 is a pulse-response productivity model with temperature and precipitation as the primary drivers.
-The model is parameterized using observations from the Phenocam network, which have a small resolution  (footprints of < 1ha), sub-daily sampling and sites across all major biomes.
+The model is parameterized using observations from the PhenoCam network, which have a small spatial resolution  (footprints of < 1ha), sub-daily sampling and sites across all major biomes.
 These attributes make the PhenoGrass model potentially widely applicable.
-We expand on the evaluation of the original study by using 445 additional site-years of data from 77 new sites.
+We expand on the evaluation of the original study by using 84 PhenoCam sites, totalling 89 distinct time series, with 463 site-years of data.
 We test the model’s performance across varying combinations of North American ecoregions and vegetation types to find an optimal spatial extent in which to parameterize and apply the model.
@@
 Model inputs are daily precipitation, temperature, potential evapotranspiration (derived from the Hargreaves equation, @hargreaves1985), and solar radiation.
-The primary output is fractional vegetation cover.
+The primary output is fractional vegetation cover (fCover).
 The original model form, derived in Choler et 2010 and Choler et al.
 2011, used only temperature and potential evapotranspiration and was parameterized using satellite-derived NDVI data.
-@hufkens2016 expanded on the original Choler model by incorporating growth and senescence restraints from temperature and solar radiation, and also included a scaling factor to convert Phenocam G\textsubscript{cc} data to a fractional cover estimate.
-@hufkens2016 evaluated the PhenoGrass model using 14 grassland Phenocam sites across Western North America with a total of 34 site years.
+@hufkens2016 expanded on the original Choler model by incorporating growth and senescence restraints from temperature and solar radiation, and also included a scaling factor to convert PhenoCam G\textsubscript{cc} data to a fractional cover estimate.
+@hufkens2016 evaluated the PhenoGrass model using 14 grassland PhenoCam sites across Western North America with a total of 34 site years.
 They found the modelled fractional cover correlated well with annual productivity at both a daily and annual timescale.
@@
 
-The Phenocam network is a global network of fixed, near-surface cameras capturing true-color images of vegetation throughout the day [@richardson2018a].
+The PhenoCam network is a global network of fixed, near-surface cameras capturing true-color images of vegetation throughout the day [@richardson2018a].
 Using a ratio of the three RGB bands a greenness metric (green chromatic coordinate, G\textsubscript{cc}) is calculated from each image, resulting in a daily scale time series of canopy greeness.
@@
 
-We downloaded all Phenocam data with ROIs of the grasslands (GR), shrublands (SH), and agricultural (AG) vegetation types for the years 2012 to 2018, totalling 91 sites and 479 site-years (Fig.
+We downloaded all Phenocam data with ROIs of the grasslands (GR), shrublands (SH), and agricultural (AG) vegetation types for the years 2012 to 2018, totalling 89 distinct time series  and 463 site-years (Fig.
 1, Table A1).
@@
 
-We used a three-step process to evaluate the PhenoGrass model at different spatial extents.
-First, we parameterized a suite of 15 PhenoGrass models comprising four different extents to find the most appropriate one to make forecasts (Fig.
+To find the most appropriate scale we evaluated the model using three different scales of vegetation type, with 11 total model parameterizations (Fig.
 2).
-The largest extent used all Phenocam locations described above (84 sites).
-Next were all sites, respectively, within the three vegetation types (grasslands, shrublands, and agricultural), and all sites, respectively, within each of the four ecoregions (North American Deserts, Great Plains, NW Forests, Eastern Temperate Forests) [@omernik2014].
-Finally, we parameterized models for each vegetation type within each ecoregion (eg.
-All grassland locations within the Great Plains ecoregion).
+The largest scale used all Phenocam locations described above (89 sites).
+Next were all sites, respectively, within the three vegetation types indicated by the ROI (grasslands, shrublands, and agricultural).
+Finally, we parameterized models for each vegetation type within each Level 1 North American Ecoregion (eg.
+All grassland sites within the Great Plains ecoregion).
 All sets of parameterized models were limited to have at least five sites.
 
-Second, using the same combinations of sites described above, we applied each model parameterization to its respective extent as well as each one below it.
-For example, sites within the grassland vegetation type were tested using the grassland only model as well as the all site  model.
-All sites within the Great Plains ecoregion were tested using the Great Plains ecoregion model as well as the all site model.
-The shrubland vegetation sites within the N.A.
-Deserts ecoregion were tested using its respective model, the N.A.
-Deserts ecoregion model, and the all site model.
-There was no cross-validation using out of sample data in this step as it would have been computationally expensive.
+We evaluated each of the 11 models using the coefficient of determination (R\textsuperscript{2}) and root mean square error (RMSE) of the daily fractional cover estimates.
+R\textsuperscript{2} and RMSE were calculated for each site and then averaged across all sites within the respective scale.
+There was no cross-validation using out of sample data in the initial fitting as it would have been computationally expensive.
 Rather, error metrics from these in-sample tests were treated as a best case scenario in what each model parameterization can achieve.
-From these results we used a threshold to select which model to evaluate further using cross-validation at the appropriate scale.
-For the error metric we chose the coefficient of determination (R\textsuperscript{2}) calculated for each site, and averaged across all sites in the respective evaluation.
-We used a threshold of 0.65, which is viewed as "acceptable" for time-series models [@ritter2013].
+From these results we used a threshold to select which models to evaluate further using cross-validation.
+The threshold value was an R\textsuperscript{2} threshold of 0.65, which is viewed as "acceptable" for time-series models [@ritter2013].
 
-```{r, out.width = "8.3cm", echo = FALSE, fig.cap = "Scaling representation of the 15 model parameterizations.
+```{r, out.width = "8.3cm", echo = FALSE, fig.cap = "Scaling representation of the 11 model parameterizations.
 Numbers in parentheses represent the number of sites and site years, respectively.
@@
 
-Third, models which exceed the threshold described above were subject to further evaluation.
-For each model we performed a leave one out cross-validation, where the model was re-fit with one Phenocam site not included in the training data, and then evaluated against the left out site.
-In this step a scaling coefficient to link mean annual precipitation with PhenoCam G\textsubscript{cc} was held constant at the value obtained in step two.
-The resulting coefficient of determination is the average among all modelled sites using their respective out of sample test.
+Models which exceed the threshold were subject to further evaluation.
+For each model we performed a leave one out cross-validation, where the model was re-fit with one Phenocam site not included in the training data, and then evaluated against this left out site.
+In this step a scaling coefficient to link mean annual precipitation with PhenoCam G\textsubscript{cc} was held constant at the value obtained in first fitting.
+The resulting R\textsuperscript{2} and RMSE are the average among all modelled sites using their respective out of sample test.
 
-All phenocam data was downloaded using the phenocamr R package [@hufkens2018].
+All phenocam data were downloaded using the phenocamr R package [@hufkens2018].
 Other packages used in the R 3.6 language were dplyr [@dplyr], tidyr [@tidyr], ggplot2 [@ggplot2], daymetr [@hufkens2018], rgdal [@rgdal2019], and sf [@sf2018].
@@
 
-At the largest extent, where the phenograss model was parameterized using all 91 sites, the model performed poorly with an average R\textsuperscript{2} value among sites of 0.33.
-The error of this largest extent model when applied across different subsets of sites (Table 1, leftmost column) shows that agricultural and North American desert sites contribute most to this low R\textsuperscript{2}, while some grassland sites performed well in this regard.
-Models built at the ecoregion or vegetation type extent also performed poorly overall.
+At the largest scale, where the PhenoGrass model was parameterized using all 89 sites, the model performed poorly with an average R\textsuperscript{2} value among sites of 0.31 (Table 1; Fig.
+3).
+Models built using all sites of a respective vegetation type performed poorly as well, though were slightly better than the all site model (Fig.
+3).
 The best model performance was achieved when models were built using a single vegetation type subset to a single ecoregion.
+Grasslands within the Great Plains and Eastern Temperate Forests ecoregions were the only instances where R\textsuperscript{2} exceeded the 0.65 threshold, though Grasslands within N.W.
+Forests came close (R\textsuperscript{2} = 0.64).
+
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "Observed and predicted fCover values of the All Site model using all available data, and the three vegetation type models each using all PhenoCam sites with an ROI in the respective vegetation type.
+"}
+knitr::include_graphics("figs/allsite_veg_errors.png")
+```
+
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "Observed and predicted fCover values for seven model iterations where only sites with a specified vegetation type within a single ecoregion were used in model fitting."}
+knitr::include_graphics("figs/ecoregion_vegtype_errors.png")
+```
+
+In all 11 iterations the PhenoGrass model tended to underestimate the highest fCover values, and to a lesser degree overpredict the lowest values (Figs.
+3,4).
+The best performing iterations (Grasslands in the Great Plains and Eastern Temperate forests) minimized this effect (Fig.
+4).
+The worst performing iteration, Grasslands in N.A.
+Deserts, had little variation in predicted fCover values, resulting in the lowest R\textsuperscript{2} overall.
 
 The grassland vegetation type, subset to specific ecoregions, predominantly outperformed other iterations of the PhenoGrass model (Table 1).
-Models built using grasslands within the Eastern Temperate Forest and Great Plains ecoregions had the highest average R\textsuperscript{2} values of 0.82 and  0.70, respectively.
-Using leave one out cross-validation on these two grassland model iterations resulted in similar errors of 0.78 and 0.67 for the Eastern Temperate Forest and Great Plains, respectively.
-Grasslands in the North American deserts were not modelled well at any extent and had the lowest R\textsuperscript{2} values in the entire analysis.
+Models built using grasslands within the Eastern Temperate Forest and Great Plains ecoregions had the highest average R\textsuperscript{2} values of 0.82 and  0.69, respectively.
+Using leave one out cross-validation on these two grassland model iterations resulted in similar errors of 0.79 and 0.67 for the Eastern Temperate Forest and Great Plains, respectively.
+Though N.W.
+Forests grasslands had an in-sample R\textsuperscript{2} just below the 0.65 threshold, the cross-validation was well below it (0.52).
+Grasslands in the North American deserts were not modelled well at any scale and had the lowest R\textsuperscript{2} values in the entire analysis.
+The observed greenness patterns of these desert grasslands had extremely high variability in their magnitude and timing, with short distinct peaks in greenness and numerous off-peak fluctuations.
+The fitted model, which minimized the mean CVME among the 5 sites, was not able to reproduce this high variability and instead produced fCover values that were severely constrained to a narrow range (Fig.
+S1).
 
-Agriculture and shrubland sites were poorly modelled at all extents.
-The highest average R\textsuperscript{2} at any extent was 0.51, and 0.39 for shrubland and agriculture, respectively.
-Shrublands in North American Deserts improved as the model became more and more localized, but never exceeded the 0.65 threshold.
+Agriculture and shrubland sites were poorly modelled at all scales.
+Performance of agriculture within the E.
+Temperate Forest ecoregion (R\textsuperscript{2} = 33) improved over the All Agriculture model (R\textsuperscript{2} = 0.18), but decreased in the Great Plains (from 0.24 to 0.18).
+There was only a single ecoregion with a minimum of five shrubland sites, N.A.
+Deserts, and it performed only slightly better than the All Shrubland model.
+Shrublands in N.A.
+Deserts did not have the high variability seen in desert grasslands.
 
 \begin{table}
-\caption{Average site-level coefficient of determination (R\textsuperscript{2}) for each vegetation type and ecoregion group parameterized model applied to its respective extent as well as appropriate smaller extents.
-Bold indicates when a R\textsuperscript{2} was greater than the “acceptable” threshold of 0.65.
+\caption{Average site-level coefficient of determination (R\textsuperscript{2}) and root mean square error (RMSE) for each model parameterization.
+Bold indicates when the R\textsuperscript{2} was greater than the acceptable threshold of 0.65.
 Values in parentheses represent the average R\textsuperscript{2} in leave-1-out cross validation.}
-\begin{tabular}{l|l|l|l|l}
+\begin{tabular}{l|l|r|r|r}
 \hline
- & All Site Model & Ecoregion Model & Vegetation Model & Ecoregion+Vegetation Model\\
+ & R\textsuperscript{2} & RMSE & Num. Sites & Site Years\\
 \hline
-All Sites & 0.33 &  &  & \\
+All Sites & 0.31 & 0.14 & 89 & 462.5\\
 \hline
-\textbf{E. Temperate Forests} & 0.43 & 0.48 &  & \\
+All Agriculture & 0.24 & 0.19 & 38 & 175.7\\
 \hline
-\hspace{1em}Agriculture & 0.34 & 0.39 & 0.38 & 0.33\\
+All Grasslands & 0.45 & 0.16 & 37 & 205.8\\
 \hline
-\hspace{1em}Grasslands & \textbf{0.73} & \textbf{0.79} & \textbf{0.65} & \textbf{0.82 (0.78)}\\
+All Shrublands & 0.44 & 0.09 & 14 & 81.0\\
 \hline
-\textbf{Great Plains} & 0.45 & 0.45 &  & \\
+\textbf{E. Temperate Forests} &  &  &  & \\
 \hline
-\hspace{1em}Agriculture & 0.22 & 0.15 & 0.17 & 0.18\\
+\hspace{1em}Agriculture & 0.33 & 0.12 & 20 & 99.2\\
 \hline
-\hspace{1em}Grasslands & 0.64 & \textbf{0.70} & \textbf{0.67} & \textbf{0.70 (0.67)}\\
+\hspace{1em}Grasslands & \textbf{0.82 (0.79)} & 0.07 & 6 & 35.7\\
 \hline
-\textbf{N. American Deserts} & -0.11 & 0.14 &  & \\
+\textbf{Great Plains} &  &  &  & \\
 \hline
-\hspace{1em}Grasslands & -0.51 & -0.22 & -0.51 & -0.08\\
+\hspace{1em}Agriculture & 0.18 & 0.19 & 13 & 54.3\\
 \hline
-\hspace{1em}Shrublands & 0.15 & 0.37 & 0.40 & 0.51\\
+\hspace{1em}Grasslands & \textbf{0.69 (0.67)} & 0.13 & 15 & 71.5\\
 \hline
-\textbf{N.W. Forests} & 0.25 & 0.58 &  & \\
+\textbf{N. American Deserts} &  &  &  & \\
 \hline
-\hspace{1em}Grasslands & 0.30 & 0.61 & 0.36 & 0.64\\
+\hspace{1em}Grasslands & -0.04 & 0.19 & 5 & 20.3\\
 \hline
-\textbf{All Agriculture} & 0.27 &  & 0.26 & \\
+\hspace{1em}\textbf{Shrublands} & 0.52 & 0.16 & 8 & 39.7\\
 \hline
-\textbf{All Grasslands} & 0.43 &  & 0.45 & \\
+\textbf{N.W. Forests} &  &  &  & \\
 \hline
-\textbf{All Shrublands} & 0.23 &  & 0.39 & \\
+\hspace{1em}\textbf{Grasslands} & 0.64 (0.52) & 0.08 & 7 & 50.7\\
 \hline
@@
 
-We performed an extensive evaluation of the PhenoGrass model across ecoregions and vegetation types to determine the best extent at which to parameterize and apply the model.
+We performed an extensive evaluation of the PhenoGrass model across ecoregions and vegetation types to determine the best scale at which to parameterize and apply the model.
 We found the model most suitable to grassland vegetation when constrained to the ecoregion level, though it did not perform well in grasslands in the North American desert ecoregion.
-Shrublands and agriculture were not well represented by the model regardless of the extent.
+Shrublands and agriculture were not well represented by the model regardless of the scale.
 Results from this study will facilitate long-term forecasts of grassland productivity constrained to an appropriate vegetation type and extent.
@@
 Grasslands in the N.A.
-Desert biome coexist with shrubs, resulting in complex water usage dynamics described in more detail below.
+Desert biome coexist with shrubs, resulting in complex water use dynamics described in more detail below.
 The pulse-response design of PhenoGrass, which makes it well suited in areas with high cover of perennial grass, is likely not applicable when grasses are interspersed with woody plants.
 
-Shrublands were not well modelled at any extent.
-Dryland shrubs, representing 8 of the 15 shrubland Phenocams analysed here, coexist with grasses by accessing different pools of soil water [@weltzin2000; @muldavin2008], thus have different responses to precipitation and resulting greenness patterns [@browning2017; @yan2019].
-A prior form of the PhenoGrass model was designed to work with dryland shrubs [@ogle2004], yet here PhenoGrass showed little effectiveness.
-The current model form has only a single soil water pool with fluxes from precipitation and evapotranspiration, thus is not well-suited for representing the deeper water pools which shrubs use.
+Shrublands were not well modelled at any scale.
+Dryland shrubs, representing 8 of the 15 shrubland PhenoCams analysed here, coexist with grasses by accessing different pools of soil water [@weltzin2000; @muldavin2008], thus have different responses to precipitation and resulting greenness patterns [@browning2017; @yan2019].
+A prior form of the PhenoGrass model was designed to work with dryland shrubs by using two soil water pools [@ogle2004], yet here PhenoGrass, with a single soil water pool, was less effective for shrubland vegetation.
+The single pool of the PhenoGrass model is coupled with fluxes from precipitation and evapotranspiration, thus is not well-suited for representing the deeper water pools that shrubs can routinely access[@shenk2002; @ward2013].
 Potential improvements would likely need to incorporate a deep soil water pool, in addition to the shallow, which are each utilized by the respective plant functional groups.
@@
 Management practices of crops artificially increase productivity beyond what would naturally occur, and planting and harvest result in abrupt changes in greenness metrics [@begue2018].
-While the results here are unsurprising, to our knowledge this is the first attempt to use near-surface images to drive an agricultural productivity model.
-We have shown that the PhenoGrass model, designed for natural systems, does not generalize to managed agriculture crops.
-Future work in using Phenocam data to model agricultural productivity would likely need to incorporate crop specific parameters and management activity, which other cropland modelling systems use [@fritz2019].
-The integration of the Phenocam network within the Long-Term Agricultural Research (LTAR) will likely be beneficial for this, as the timing and intensity of management activities or experimental treatments can be incorporated into modelling efforts.
+While the results were not necessarily surprising, to our knowledge this is the first attempt to use near-surface images to drive an productivity model for agricultural vegetation.
+We have shown that the PhenoGrass model, designed for natural systems, does not generalize to actively managed agricultural systems.
+Future work in using PhenoCam data to model agricultural productivity would likely need to incorporate crop specific parameters and management activity, which other cropland modelling systems use [@fritz2019].
+The integration of the PhenoCam network within the Long-Term Agricultural Research (LTAR) will likely be beneficial for this, as the timing and intensity of management activities or experimental treatments can be incorporated into modelling efforts.
 
-@hufkens2016 originally evaluated the Phenograss model using 14 grassland sites distributed among seven North American ecoregions.
+@hufkens2016 originally evaluated the PhenoGrass model using 14 grassland sites distributed among seven North American ecoregions.
 In their evaluation they had an average R\textsuperscript{2} of 0.71, while here the model performed poorly when using more than 1 ecoregion.
 It’s likely that the original 14 grassland sites were ideal locations for the PhenoGrass model, since on average they have a single greenup season every year in the spring or summer (Fig.
-3A).
+5A).
 The additional 24 grassland sites used in the current study have high seasonal variability and elongated growing seasons (Fig.
-3B), and were thus more difficult to represent in a single continental scale grassland model.
+5B, Fig.
+S1), and were thus more difficult to represent in a single continental scale grassland model.
 
@@
 Each line represents the long-term average Green Chromatic Coordinate of a single site across all available years, smoothed using a GAM model."}
-knitr::include_graphics("figs/fig3_discussion_avg_site_gcc.png")
+knitr::include_graphics("figs/fig5_discussion_avg_site_gcc.png")
 ```

### 2020-08-27 analysis/discussion_figure_site_gcc_comparison.R
knock the figure dpi down a bit

@@ -32,7 +32,7 @@ discusion_figure = ggplot(grassland_gcc, aes(x=doy, y=gcc, color=as.factor(times
         strip.text = element_text(hjust = 0,size=16)) +
   labs(x='Day of Year', y='Avg. GCC')
 
-ggsave('manuscript/figs/fig5_discussion_avg_site_gcc.png', plot=discusion_figure, height = 15, width=30, units = 'cm', dpi=200)  
+ggsave('manuscript/figs/fig5_discussion_avg_site_gcc.png', plot=discusion_figure, height = 15, width=30, units = 'cm', dpi=150)  
   
 
   # Also while where here parse out what the original ecoregions were

### 2020-08-27 manuscript/manuscript.Rmd
fix table boldness
Manuscript updates 1st revission (#10)
revisions for "resubmission" to biogeosciences.
new figures too

@@
 Performance of agriculture within the E.
-Temperate Forest ecoregion (R\textsuperscript{2} = 33) improved over the All Agriculture model (R\textsuperscript{2} = 0.18), but decreased in the Great Plains (from 0.24 to 0.18).
+Temperate Forest ecoregion (R\textsuperscript{2} = 0.33) improved over the All Agriculture model (R\textsuperscript{2} = 0.18), but decreased in the Great Plains (from 0.24 to 0.18).
 There was only a single ecoregion with a minimum of five shrubland sites, N.A.
@@
 \hline
-\hspace{1em}\textbf{Shrublands} & 0.52 & 0.16 & 8 & 39.7\\
+\hspace{1em}Shrublands & 0.52 & 0.16 & 8 & 39.7\\
 \hline
@@
 \hline
-\hspace{1em}\textbf{Grasslands} & 0.64 (0.52) & 0.08 & 7 & 50.7\\
+\hspace{1em}Grasslands & 0.64 (0.52) & 0.08 & 7 & 50.7\\
 \hline
@@
 A prior form of the PhenoGrass model was designed to work with dryland shrubs by using two soil water pools [@ogle2004], yet here PhenoGrass, with a single soil water pool, was less effective for shrubland vegetation.
-The single pool of the PhenoGrass model is coupled with fluxes from precipitation and evapotranspiration, thus is not well-suited for representing the deeper water pools that shrubs can routinely access[@shenk2002; @ward2013].
+The single pool of the PhenoGrass model is coupled with fluxes from precipitation and evapotranspiration, thus is not well-suited for representing the deeper water pools that shrubs can routinely access [@schenk2002; @ward2013].
 Potential improvements would likely need to incorporate a deep soil water pool, in addition to the shallow, which are each utilized by the respective plant functional groups.

### 2020-08-27 analysis/discussion_figure_site_gcc_comparison.R

@@ -32,7 +32,7 @@ discusion_figure = ggplot(grassland_gcc, aes(x=doy, y=gcc, color=as.factor(times
         strip.text = element_text(hjust = 0,size=16)) +
   labs(x='Day of Year', y='Avg. GCC')
 
-ggsave('manuscript/figs/fig3_discussion_avg_site_gcc.png', plot=discusion_figure, height = 15, width=30, units = 'cm', dpi=200)  
+ggsave('manuscript/figs/fig5_discussion_avg_site_gcc.png', plot=discusion_figure, height = 15, width=30, units = 'cm', dpi=150)  
   
 
   # Also while where here parse out what the original ecoregions were

### 2020-08-27 analysis/error_figures.R

@@ -65,7 +65,7 @@ allsite_veg_error_fig = ggplot(allsite_veg_predictions, aes(x=fCover_observed, y
         strip.text = element_text(hjust = 0, size=10)) +
   labs(x='Observed fCover', y='Predicted fCover') 
 
-ggsave('manuscript/figs/allsite_veg_errors.png',plot=allsite_veg_error_fig, height=12, width = 12, units='cm', dpi=200)
+ggsave('manuscript/figs/allsite_veg_errors.png',plot=allsite_veg_error_fig, height=12, width = 12, units='cm', dpi=150)
 
 
 #######################################################
@@ -110,4 +110,4 @@ ecoregion_vegtype_error_figure = ggplot(ecoregion_vegtype_predictions, aes(x=fCo
         strip.text = element_text(hjust = 0, size=8)) +
   labs(x='Observed fCover', y='Predicted fCover') 
 
-ggsave('manuscript/figs/ecoregion_vegtype_errors.png',plot=ecoregion_vegtype_error_figure, height=20, width = 15, units='cm', dpi=200)
+ggsave('manuscript/figs/ecoregion_vegtype_errors.png',plot=ecoregion_vegtype_error_figure, height=20, width = 15, units='cm', dpi=150)

### 2020-08-27 analysis/make_error_tables.R

@@ -29,7 +29,7 @@ error_table_order = tribble(
 
 )
 
-bold_row_titles = c(5,8,11,13,14,15)
+bold_row_titles = c(5,8,11,14)
 
 error_table = primary_errors %>%
   full_join(error_table_order, by=c('fitting_set'='model_fitting_sets')) %>%

### 2020-12-27 manuscript/manuscript.Rmd
appendex- > supp
small revisions
- forecast -> projections
- extent -> scale
- add "spatial scale" clarification

@@
 abstract: |
-  Grasslands provide many important ecosystem services globally and forecasting grassland productivity in the coming decades will provide valuable information to land managers.
-Productivity models can be well-calibrated at local scales, but generally have some maximum spatial extent in which they perform well.
-Here we evaluate a grassland productivity model to find the optimal spatial extent for parameterization, and thus for subsequently applying it in future forecasts for North America.
+  Grasslands provide many important ecosystem services globally and projecting grassland productivity in the coming decades will provide valuable information to land managers.
+Productivity models can be well-calibrated at local scales, but generally have some maximum spatial scale in which they perform well.
+Here we evaluate a grassland productivity model to find the optimal spatial scale for parameterization, and thus for subsequently applying it in future productivity projections for North America.
 We also evaluated the model on new vegetation types to ascertain its potential generality.
@@
 It also performed poorly in agriculture vegetation, likely due to management activities, and shrubland vegetation, likely because the model lacks representation of deep water pools.
-This work allows us to perform long-term forecasts in areas where model performance has been verified, with gaps filled in by future modelling efforts.
+This work allows us to perform long-term projections in areas where model performance has been verified, with gaps filled in by future modelling efforts.
 
@@
 For example, even with consistent shifts in climate, different locations can experience different changes in productivity due to local-scale responses [@zhang2011;@sala2012;@knapp2017].
-This highlights the need for models which can be resolved at small spatial and temporal scales, thus making long-term grassland productivity forecasts as informative as possible.
+This highlights the need for models which can be resolved at small spatial and temporal scales, thus making long-term grassland productivity projections as informative as possible.
 
-There are several potential limitations in the underlying productivity models which can drive such a forecast.
-Process-based models parameterized with observed data have limited transferability beyond the spatial extent from which their training data came [@taylor2019a].
+There are several potential limitations in the underlying productivity models which can drive such a projection.
+Process-based models parameterized with observed data have limited transferability beyond the spatial scale from which their training data came [@taylor2019a].
 For any location the most accurate model will be one which was parameterized from locally collected data, yet these site-specific models will not generalize to new locations [@basler2016].
 Incorporating more, and diverse, locations into the model building process will allow it to be more generalizable, yet this comes at a cost of decreased proficiency at all locations [@garcia-mozo2008; @basler2016].
-Thus there is an optimal extent in the building and subsequent application of productivity models, which depends on a tradeoff between proficiency at the local scale and applicability at the larger scale.
+Thus there is an optimal scale in the building and subsequent application of productivity models, which depends on a tradeoff between proficiency at the local scale and applicability at the larger scale.
 
-Here we evaluate a productivity model with the intention of it driving long-term forecasts.
+Here we evaluate a productivity model with the intention of it driving long-term projections.
 The PhenoGrass model developed by @hufkens2016 is a pulse-response productivity model with temperature and precipitation as the primary drivers.
@@
 We expand on the evaluation of the original study by using 84 PhenoCam sites, totalling 89 distinct time series, with 463 site-years of data.
-We test the model’s performance across varying combinations of North American ecoregions and vegetation types to find an optimal spatial extent in which to parameterize and apply the model.
-Finally we address where the model performs poorly and how productivity forecasts for these areas could be implemented or improved.
+We test the model’s performance across varying combinations of North American ecoregions and vegetation types to find an optimal spatial scale in which to parameterize and apply the model.
+Finally we address where the model performs poorly and how productivity projections for these areas could be implemented or improved.
 
@@
 We downloaded all Phenocam data with ROIs of the grasslands (GR), shrublands (SH), and agricultural (AG) vegetation types for the years 2012 to 2018, totalling 89 distinct time series  and 463 site-years (Fig.
-1, Table A1).
+1, Table S1).
 As input to the PhenoGrass model we used the 3-day smoothed G\textsubscript{cc} scaled, for each ROI, from 0-1.
@@
 
-To find the most appropriate scale we evaluated the model using three different scales of vegetation type, with 11 total model parameterizations (Fig.
+To find the most appropriate scale we evaluated the model using three different combinations of ecoregion and vegetation type, with 11 total model parameterizations (Fig.
 2).
+Here we use the term “spatial scale” to refer to the combination of ecoregion/s and vegetation type/s used within each model.
+This includes using all vegetation types with an ecoregion, or all sites of a specific vegetation type from several ecoregions.
 The largest scale used all Phenocam locations described above (89 sites).
@@
 Shrublands and agriculture were not well represented by the model regardless of the scale.
-Results from this study will facilitate long-term forecasts of grassland productivity constrained to an appropriate vegetation type and extent.
+Results from this study will facilitate long-term projections of grassland productivity constrained to an appropriate vegetation type and scale.
 
@@
 This can result in a spatial ensemble where the output for any one location and vegetation type is represented by the most appropriate model.
-Our current work will allow for long-term small scale forecasts of grassland productivity for a large fraction of North America.
+Our current work will allow for long-term small scale projections of grassland productivity for a large fraction of North America.

### 2020-12-27 manuscript/manuscript.Rmd
add mean cvmae equation

@@
 The threshold value was an R\textsuperscript{2} threshold of 0.65, which is viewed as "acceptable" for time-series models [@ritter2013].
+
+Parameterization was done using differential evolution, a global optimization algorithm, to minimize the mean coefficient of variation of the mean absolute error (\textit{F}, Eq.
+1), which accounts for variation among sites [@choler2011].
+
+\begin{equation}
+F = \frac{1}{N}\sum_{j=1}^{N}CVMAE_{j}
+\end{equation}
+
+\begin{equation}
+CVMAE_{j} = \frac{\frac{1}{i}\sum_{i=1}^{n}|fCover_{i,obs} - fCover_{i,pred}|}{\overline{fCover_{obs}}}
+\end{equation}
+
+where \textit{N} is the number of sites and \textit{n} is the number of daily values at each site, \textit{fCover\textsubscript{i,obs}} and \textit{fCover\textsubscript{i,pred}} are observed and predicted values, respectively.
+\textit{fCover\textsubscript{pred}} is the average observed \textit{fCover} at each site.
 

### 2020-12-27 analysis/analyze_cv_errors.R
Change error metrics in scripts, figures, & tables
R2 is now called the nash sutcliffe coeficient of efficiency, NSE,
note the math is the same

RMSE is removed since it isn't comparable across sites. replaced
by the mean coefficient of variation of the mean absolute error,
aka the mean cvmae, aka F. This is also the loss metric used in
fitting.

@@ -11,9 +11,8 @@ cv_predictions = read_csv('data/cv_model_predictions.csv',
 
 site_level_errors = cv_predictions %>%
   group_by(fitting_set, model, left_out_timeseries) %>%
-  summarise(rmse = sqrt(mean( (gcc_observed - gcc_predicted)^2 ,na.rm=T)),
-            r2   = 1  - sum((gcc_observed - gcc_predicted)^2,na.rm=T) / sum((gcc_observed - mean(gcc_observed,na.rm=T))^2,na.rm=T),
-            #mae  = mean(abs(gcc_observed - gcc_predicted),na.rm=T),
+  summarise(cvmae = mean(abs(gcc_observed - gcc_predicted), na.rm=T)/mean(gcc_observed, na.rm=T) ,
+            nse   = 1  - sum((gcc_observed - gcc_predicted)^2,na.rm=T) / sum((gcc_observed - mean(gcc_observed,na.rm=T))^2,na.rm=T),
             n=n(),
             n_timeseries = n_distinct(timeseries_id),
             percent_na=mean(is.na(gcc_observed))) %>%
@@ -24,7 +23,8 @@ assertthat::assert_that(all(site_level_errors$n_timeseries==1), msg='>1 timeseri
 
 site_level_errors %>%
   group_by(fitting_set, model) %>%
-  summarise(avg_r2 = mean(r2),
-            avg_rmse = mean(rmse)) %>%
+  summarise(avg_nse = mean(nse),
+            mean_cvmae = mean(cvmae),
+            n_sites = n()) %>%
   ungroup()
   

### 2020-12-27 analysis/make_error_tables.R

@@ -35,12 +35,12 @@ error_table = primary_errors %>%
   full_join(error_table_order, by=c('fitting_set'='model_fitting_sets')) %>%
   arrange(row_order) %>%
   mutate(row_label = cell_spec(row_label, 'latex', bold=(row_order %in% bold_row_titles))) %>%
-  mutate(r2 = cell_spec(r2, 'latex', bold=(r2>=0.65))) %>%
-  mutate(r2 = ifelse(r2=='NA',NA,r2)) %>% # cell_spec is inserting literal NA here
-  select(row_label, r2, rmse, n_timeseries, site_years)
+  mutate(nse = cell_spec(nse, 'latex', bold=(nse>=0.65))) %>%
+  mutate(nse = ifelse(nse=='NA',NA,nse)) %>% # cell_spec is inserting literal NA here
+  select(row_label, nse, mean_cvmae, n_timeseries, site_years)
   
 # TODO: need cv errors in here for the 2 grassland models
-table_column_names = c('','R\textsuperscript{2}','RMSE','Num. Sites','Site Years')
+table_column_names = c('','NSE','F','Num. Sites','Site Years')
 
 kable(error_table, 'latex', col.names = table_column_names, escape = F) %>%
   add_indent(c(6,7,9,10,12,13,15))
@@ -59,15 +59,15 @@ site_info = read_csv('site_list.csv') %>%
    left_join(ecoregion_info, by='ecoregion') %>%
    select(timeseries_id, site_label, ecoregion_abbr)
 
-# Arrange the csv file from columns c('fitting_set','model','timeseries_id','rmse','r2') 
-# to c('site_label','fitting_set','model1_r2','model1_rmse','model2_r2',...)
+# Arrange the csv file from columns c('fitting_set','model','timeseries_id','cvmae','nse') 
+# to c('site_label','fitting_set','model1_nse','model1_cvmae','model2_nse',...)
 
 site_errors = read_csv('results/site_level_error_table.csv') %>% 
    filter(timeseries_id != 32) %>% # drop jasperridge roi2000 here while the new models run.
    mutate(model = recode(model, 'NaiveMAPCorrected' = 'Naive')) %>%
    left_join(site_info, by='timeseries_id') %>%
-   select(fitting_set, model, rmse, r2, site_label, ecoregion_abbr) %>%
-   gather(error_type, error_value, r2, rmse) %>%
[…]

### 2020-12-27 manuscript/manuscript.Rmd
change errors in text
R2 is now NSE
replace RMSE with mean CVMAE (aka F)
include figure caption clarifying lines and numbers

--- 2020-12-27 manuscript/manuscript.Rmd
@@
 
-We evaluated each of the 11 models using the coefficient of determination (R\textsuperscript{2}) and root mean square error (RMSE) of the daily fractional cover estimates.
-R\textsuperscript{2} and RMSE were calculated for each site and then averaged across all sites within the respective scale.
+We evaluated each of the 11 models using the Nash Sutcliffe coefficient of efficiency (\(NSE\), Eq.
+2) as well as the mean coefficient of variation of the mean absolute error (\textit{F}, Eq.
+3) of the daily fractional cover estimates.
+\(NSE\) was calculated for each site and then averaged across all sites within the respective spatial scale (\( \overline{NSE} \), Eq.
+1).
 There was no cross-validation using out of sample data in the initial fitting as it would have been computationally expensive.
@@
 From these results we used a threshold to select which models to evaluate further using cross-validation.
-The threshold value was an R\textsuperscript{2} threshold of 0.65, which is viewed as "acceptable" for time-series models [@ritter2013].
+The threshold value was an \(\overline{NSE}\) of 0.65, which is viewed as "acceptable" for time-series models [@ritter2013].
+Parameterization was done using differential evolution, a global optimization algorithm, to minimize \textit{F}.
+
+\begin{equation}
+\overline{NSE} = \frac{1}{N}\sum_{j=1}^{N}NSE_{j}
+\end{equation}
+
+\begin{equation}
+NSE_{j} = 1- \frac{\sum_{i=1}^{n}(fCover_{obs} - fCover_{pred})^{2}}{\sum_{i=1}^{n}(fCover_{obs} - \overline{fCover_{obs}})^{2}}
+\end{equation}
+
+\begin{equation}
+F = \frac{1}{N}\sum_{j=1}^{N}CVMAE_{j}
+\end{equation}
+
+\begin{equation}
+CVMAE_{j} = \frac{\frac{1}{i}\sum_{i=1}^{n}|fCover_{i,obs} - fCover_{i,pred}|}{\overline{fCover_{obs}}}
+\end{equation}
+
+where \textit{N} is the number of sites within the spatial scale evaluated and \textit{n} is the number of daily values at each site, \textit{fCover\textsubscript{i,obs}} and \textit{fCover\textsubscript{i,pred}} are observed and predicted values, respectively.
+\textit{fCover\textsubscript{pred}} is the average observed \textit{fCover} at each site.
 
@@
 In this step a scaling coefficient to link mean annual precipitation with PhenoCam G\textsubscript{cc} was held constant at the value obtained in first fitting.
-The resulting R\textsuperscript{2} and RMSE are the average among all modelled sites using their respective out of sample test.
+The resulting \(\overline{NSE}\) and \textit{F} are from all modelled sites using their respective out of sample test.
 
@@
 
-At the largest scale, where the PhenoGrass model was parameterized using all 89 sites, the model performed poorly with an average R\textsuperscript{2} value among sites of 0.31 (Table 1; Fig.
+At the largest scale, where the PhenoGrass model was parameterized using all 89 sites, the model performed poorly with an \(\overline{NSE}\) value among sites of 0.31 (Table 1; Fig.
 3).
@@
 The best model performance was achieved when models were built using a single vegetation type subset to a single ecoregion.
-Grasslands within the Great Plains and Eastern Temperate Forests ecoregions were the only instances where R\textsuperscript{2} exceeded the 0.65 threshold, though Grasslands within N.W.
-Forests came close (R\textsuperscript{2} = 0.64).
+Grasslands within the Great Plains and Eastern Temperate Forests ecoregions were the only instances where \(\overline{NSE}\) exceeded the 0.65 threshold, though Grasslands within N.W.
+Forests came close (NSE = 0.64).
 
 ```{r, out.width = "12cm", echo = FALSE, fig.cap = "Observed and predicted fCover values of the All Site model using all available data, and the three vegetation type models each using all PhenoCam sites with an ROI in the respective vegetation type.
-"}
+NSE is the average Nash Sutcliffe coefficient of efficiency among sites, and \textit{F} is the mean coefficient of variation of the mean absolute error.
+The correlation line (blue) represents the overall trend in predicted versus observed values, while the 1:1 line (red) represents a perfect fit."}
 knitr::include_graphics("figs/allsite_veg_errors.png")
@@
 
-```{r, out.width = "12cm", echo = FALSE, fig.cap = "Observed and predicted fCover values for seven model iterations where only sites with a specified vegetation type within a single ecoregion were used in model fitting."}
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "Observed and predicted fCover values for seven model iterations where only sites with a specified vegetation type within a single ecoregion were used in model fitting.
+NSE is the average Nash Sutcliffe coefficient of efficiency among sites, and \textit{F} is the mean coefficient of variation of the mean absolute error.
+The correlation line (blue) represents the overall trend in predicted versus observed values, while the 1:1 line (red) represents a perfect fit."}
 knitr::include_graphics("figs/ecoregion_vegtype_errors.png")
@@
 The worst performing iteration, Grasslands in N.A.
-Deserts, had little variation in predicted fCover values, resulting in the lowest R\textsuperscript{2} overall.
+Deserts, had little variation in predicted fCover values, resulting in the lowest \(\overline{NSE}\) overall.
 
 The grassland vegetation type, subset to specific ecoregions, predominantly outperformed other iterations of the PhenoGrass model (Table 1).
-Models built using grasslands within the Eastern Temperate Forest and Great Plains ecoregions had the highest average R\textsuperscript{2} values of 0.82 and  0.69, respectively.
+Models built using grasslands within the Eastern Temperate Forest and Great Plains ecoregions had the highest \(\overline{NSE}\) values of 0.82 and  0.69, respectively.
 Using leave one out cross-validation on these two grassland model iterations resulted in similar errors of 0.79 and 0.67 for the Eastern Temperate Forest and Great Plains, respectively.
 Though N.W.
-Forests grasslands had an in-sample R\textsuperscript{2} just below the 0.65 threshold, the cross-validation was well below it (0.52).
-Grasslands in the North American deserts were not modelled well at any scale and had the lowest R\textsuperscript{2} values in the entire analysis.
+Forests grasslands had an in-sample \(\overline{NSE}\) just below the 0.65 threshold, the cross-validation was well below it (0.52).
+Grasslands in the North American deserts were not modelled well at any scale and had the lowest \(\overline{NSE}\) values in the entire analysis.
 The observed greenness patterns of these desert grasslands had extremely high variability in their magnitude and timing, with short distinct peaks in greenness and numerous off-peak fluctuations.
@@
 Performance of agriculture within the E.
-Temperate Forest ecoregion (R\textsuperscript{2} = 0.33) improved over the All Agriculture model (R\textsuperscript{2} = 0.18), but decreased in the Great Plains (from 0.24 to 0.18).
+Temperate Forest ecoregion (NSE = 0.33) improved over the All Agriculture model (NSE = 0.18), but decreased in the Great Plains (from 0.24 to 0.18).
 There was only a single ecoregion with a minimum of five shrubland sites, N.A.
@@
 \begin{table}
-\caption{Average site-level coefficient of determination (R\textsuperscript{2}) and root mean square error (RMSE) for each model parameterization.
-Bold indicates when the R\textsuperscript{2} was greater than the acceptable threshold of 0.65.
-Values in parentheses represent the average R\textsuperscript{2} in leave-1-out cross validation.}
+\caption{Average site-level Nash Sutcliffe coefficient of efficiency (\(\overline{NSE}\)) and mean coefficient of variation of the mean absolute error (\textit{F}) for each model parameterization.
+Bold indicates when the \(\overline{NSE}\) was greater than the acceptable threshold of 0.65.
+Values in parentheses represent the \(\overline{NSE}\) or \textit{F} in leave-1-out cross validation.}
 \begin{tabular}{l|l|r|r|r}
 \hline
- & R\textsuperscript{2} & RMSE & Num. Sites & Site Years\\
+& \(\overline{NSE}\) & \textit{F} & Num. Sites & Site Years\\
 \hline
-All Sites & 0.31 & 0.14 & 89 & 462.5\\
+All Sites & 0.31 & 0.50 & 89 & 462.5\\
 \hline
-All Agriculture & 0.24 & 0.19 & 38 & 175.7\\
+All Agriculture & 0.24 & 0.51 & 38 & 175.7\\
 \hline
-All Grasslands & 0.45 & 0.16 & 37 & 205.8\\
+All Grasslands & 0.45 & 0.44 & 37 & 205.8\\
 \hline
-All Shrublands & 0.44 & 0.09 & 14 & 81.0\\
+All Shrublands & 0.44 & 0.40 & 14 & 81.0\\
 \hline
@@
 \hline
-\hspace{1em}Agriculture & 0.33 & 0.12 & 20 & 99.2\\
+\hspace{1em}Agriculture & 0.33 & 0.45 & 20 & 99.2\\
 \hline
-\hspace{1em}Grasslands & \textbf{0.82 (0.79)} & 0.07 & 6 & 35.7\\
+\hspace{1em}Grasslands & \textbf{0.82 (0.79)} & 0.28 (0.30) & 6 & 35.7\\
 \hline
@@
 \hline
-\hspace{1em}Agriculture & 0.18 & 0.19 & 13 & 54.3\\
+\hspace{1em}Agriculture & 0.18 & 0.57 & 13 & 54.3\\
 \hline
-\hspace{1em}Grasslands & \textbf{0.69 (0.67)} & 0.13 & 15 & 71.5\\
+\hspace{1em}Grasslands & \textbf{0.69 (0.67)} & 0.39 (0.40) & 15 & 71.5\\
 \hline
@@
 \hline
-\hspace{1em}Grasslands & -0.04 & 0.19 & 5 & 20.3\\
+\hspace{1em}Grasslands & -0.04 & 0.44 & 5 & 20.3\\
 \hline
-\hspace{1em}Shrublands & 0.52 & 0.16 & 8 & 39.7\\
+\hspace{1em}Shrublands & 0.52 & 0.33 & 8 & 39.7\\
 \hline
@@
 \hline
-\hspace{1em}Grasslands & 0.64 (0.52) & 0.08 & 7 & 50.7\\
+\hspace{1em}Grasslands & 0.64 & 0.34 & 7 & 50.7\\
 \hline

### 2020-12-27 manuscript/manuscript.Rmd
new 2nd paragraph in intro

@@
 
-There are several potential limitations in the underlying productivity models which can drive such a projection.
-Process-based models parameterized with observed data have limited transferability beyond the spatial scale from which their training data came [@taylor2019a].
-For any location the most accurate model will be one which was parameterized from locally collected data, yet these site-specific models will not generalize to new locations [@basler2016].
-Incorporating more, and diverse, locations into the model building process will allow it to be more generalizable, yet this comes at a cost of decreased proficiency at all locations [@garcia-mozo2008; @basler2016].
-Thus there is an optimal scale in the building and subsequent application of productivity models, which depends on a tradeoff between proficiency at the local scale and applicability at the larger scale.
+A promising method is low dimensional models, which are process models with some simplified components [@choler2010; @choler2011].
+For example, a low dimensional model might approximate transpiration to a function of potential evapotranspiration, soil available water, and live vegetation cover along with a single parameter.
+As opposed to a high dimensional model with multiple functions accounting for leaf area index, stomatal conductance, rooting depth and surface area, etc [@caylor2009; @asbjornsen2011].
+The low dimensional model is advantageous since it can generalize across broad regions with relatively few inputs.
+Yet they are still susceptible to over-fitting to local conditions since parameters or model structure can be tied to specific locations or plant functional groups [@fisher2020].
+Thus parameterizing low-dimensional models must be done with care such that they are applicable to a broad area while maintaining an acceptable level of accuracy.
 
-Here we evaluate a productivity model with the intention of it driving long-term projections.
+Here we evaluate a low-dimensional model with the intention of it driving productivity projections.
 The PhenoGrass model developed by @hufkens2016 is a pulse-response productivity model with temperature and precipitation as the primary drivers.

### 2020-12-27 manuscript/manuscript.Rmd
add text on using phenograss beyond grass

@@
 They found the modelled fractional cover correlated well with annual productivity at both a daily and annual timescale.
+Despite it's name the PhenoGrass model can theoretically apply to any vegetation type with a distinct growth signal in response to precipitation, as hypothesized in the original threshold-delay model [@ogl2004].
+Here we use two additional vegetation types, shrubs and agricultural plots, to test how applicable it is beyond grasslands.
 

### 2020-12-27 manuscript/manuscript.Rmd
fix daymet nouns

--- 2020-12-27 manuscript/manuscript.Rmd
@@
 They found the modelled fractional cover correlated well with annual productivity at both a daily and annual timescale.
+Despite it's name the PhenoGrass model can theoretically apply to any vegetation type with a distinct growth signal in response to precipitation, as hypothesized in the original threshold-delay model [@ogle2004].
+Here we use two additional vegetation types, shrubs and agricultural plots, to test how applicable it is beyond grasslands.
 
@@
 
-For historic precipitation and temperature we used the daily 4-km resolution Daymet dataset [@daymet].
-Climate time series were extracted for the pixel at the location of each phenocam tower.
+For historic precipitation and temperature we used the 1-km resolution Daymet dataset [@daymet], extracting daily time series for the pixel at each PhenoCam tower location.
 Daily mean temperature was calculated as the average between the Daymet daily minimum and maximum temperature, and smoothed with a 15 day moving average.

### 2020-12-27 manuscript/manuscript.Rmd
clarify text in fig 3/4 legends

@@
 
-```{r, out.width = "12cm", echo = FALSE, fig.cap = "Observed and predicted fCover values of the All Site model using all available data, and the three vegetation type models each using all PhenoCam sites with an ROI in the respective vegetation type.
-NSE is the average Nash Sutcliffe coefficient of efficiency among sites, and \textit{F} is the mean coefficient of variation of the mean absolute error.
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "Observed and predicted daily fCover values of the All Site model and the three vegetation type models, each using all available sites and years with the respective spatial scale.
+NSE is the average Nash Sutcliffe coefficient of efficiency among sites, and \\textit{F} is the mean coefficient of variation of the mean absolute error.
 The correlation line (blue) represents the overall trend in predicted versus observed values, while the 1:1 line (red) represents a perfect fit."}
@@
 
-```{r, out.width = "12cm", echo = FALSE, fig.cap = "Observed and predicted fCover values for seven model iterations where only sites with a specified vegetation type within a single ecoregion were used in model fitting.
-NSE is the average Nash Sutcliffe coefficient of efficiency among sites, and \textit{F} is the mean coefficient of variation of the mean absolute error.
+```{r, out.width = "12cm", echo = FALSE, fig.cap = "Observed and predicted daily fCover values for models from seven spatial scales, where only specific vegetation types within a single ecoregion were used in model fitting.
+Each uses all available sites and years with the respective spatial scale.
+NSE is the average Nash Sutcliffe coefficient of efficiency among sites, and \\textit{F} is the mean coefficient of variation of the mean absolute error.
 The correlation line (blue) represents the overall trend in predicted versus observed values, while the 1:1 line (red) represents a perfect fit."}

### 2020-12-27 manuscript/manuscript.Rmd
eq corrections, add note for new supp

@@
 2011, used only temperature and potential evapotranspiration and was parameterized using satellite-derived NDVI data.
-@hufkens2016 expanded on the original Choler model by incorporating growth and senescence restraints from temperature and solar radiation, and also included a scaling factor to convert PhenoCam G\textsubscript{cc} data to a fractional cover estimate.
+@hufkens2016 expanded on the original Choler model by incorporating growth and senescence restraints from temperature and solar radiation, and also included a scaling factor to convert PhenoCam G\textsubscript{cc} data to a fractional cover estimate (see supplemental equations).
 @hufkens2016 evaluated the PhenoGrass model using 14 grassland PhenoCam sites across Western North America with a total of 34 site years.
@@
 \begin{equation}
-CVMAE_{j} = \frac{\frac{1}{i}\sum_{i=1}^{n}|fCover_{i,obs} - fCover_{i,pred}|}{\overline{fCover_{obs}}}
+CVMAE_{j} = \frac{\frac{1}{n}\sum_{i=1}^{n}|fCover_{i,obs} - fCover_{i,pred}|}{\overline{fCover_{obs}}}
 \end{equation}
@@
 where \textit{N} is the number of sites within the spatial scale evaluated and \textit{n} is the number of daily values at each site, \textit{fCover\textsubscript{i,obs}} and \textit{fCover\textsubscript{i,pred}} are observed and predicted values, respectively.
-\textit{fCover\textsubscript{pred}} is the average observed \textit{fCover} at each site.
+\(\overline{fCover_{obs}}\) is the average observed \textit{fCover} at each site.
 

### 2020-12-27 manuscript/supplement2.Rmd
new supplement 2
all the equations and parameter descriptions.

---
output: pdf_document
---

## Multi-scale assessment of a grassland productivity model
**Shawn D. Taylor, Dawn M. Browning**

Supplementary Model Equations and Parameters

The PhenoGrass model from Hufkens et al. 2016 takes the following form:

\begin{equation}
W_{t+1} = W_{t} + P_{t} - (1-V_{t})/(D_{t}/(W_{cap}-{W_p}))^{2}E_{t}-gb_{3}D_{t}V_{t}
\end{equation}

\begin{equation}
V_{t+1} = V_{t} + gS_{t}b_{1}D_{t-L}(1-V_{t}/V_{max})-db_{2}(1-V_{t})V_{t}
\end{equation}

\begin{equation}
D_{t} = max(0,W_{t}-W_{p})
\end{equation}

\begin{equation}
W_{t} = max(0,min(W_{cap},W_{t}))
\end{equation}

\begin{equation}
V_{t} = max(0,min(V_{max},V_{t}))
\end{equation}

\begin{equation}
\left\{\begin{matrix}
D_{t-L} > D_{t-L-1}; & d=0\\ 
D_{t-L} \leq  D_{t-L-1}; & d=1 
\end{matrix}\right.
\end{equation}

\begin{equation}
S_{t} = (TOA_{t} - Phmin)/(Phmax - Phmin)
\end{equation}

\begin{equation}
\left\{\begin{matrix}
S_{t-1} > S_{t}; & d=1  \\
S_{t-1} \leq  S_{t}; & d=0 
\end{matrix}\right.
\end{equation}

\begin{equation}
g = (\frac{T_{max}-T_{m}}{T_{max}-T_{opt}})(\frac{T_{m}}{T_{opt}})^{(\frac{T_{opt}}{T_{max}-T_{opt}})}
\end{equation}

The vegetaion parameter V corresponds to the fCover, which is mapped to PhenoCam G\textsubscript{cc} using the transfer function:

\begin{equation}
S_{c} = MAP/(MAP+h)
\end{equation}

where:
\begin{equation}
fCover = G_{cc}*S_{c}
\end{equation}

MAP is the site level mean annual precipitation, and \textit{h} is an estimated parameter. G\textsubscript{cc} is a ratio of the green band within the daily image ROI:

\begin{equation}
G_{cc} = \frac{G_{dn}}{G_{dn}+R_{dn}+B_{dn}}
\end{equation}

Where G\textsubscript{dn}, R\textsubscript{dn}, and B\textsubscript{dn} are the mean digitial numbers for green, red, and blue channels, respectively. See Richardson et al. 2018 for further G\textsubscript{cc} processing details.

\newpage

\begin{table}[]
\begin{tabular}{lll}
                              & Parameter               & Description                                  \\
\textbf{State Variables}      & W                       & Soil Water Content (mm)                      \\
                             & V                       & Fractional Cover (fCover)                    \\
                             & D                       & Plant available water (mm)                   \\
                             & g                       & Plant temperature response function          \\
\textbf{Input Variables}      & P                       & Precipitaiton (mm)                           \\
                             & MAP                     & Mean annual precipitation (mm)                \\
                             & W\textsubscript{cap}    & Soil Field Capacity (mm)                     \\
                             & W\textsubscript{p}      & Soil Wilting Point (mm)                      \\
                             & V\textsubscript{max}    & Maximum vegetation cover (1.0)               \\
                             & E                       & Evapotranspiration (mm)                      \\
                             & T\textsubscript{m}      & Running 15 day mean temperature (C)          \\
                             & TOA                     & Top of atmosphere radiation                  \\
                             & T\textsubscript{max}    & Maximum temperature for growth (45 C)        \\
                             & T\textsubscript{min}    & Minimum temperature for growth (0 C)         \\
\textbf{Estimated Parameters} & b1                      & Vegetation growth parameter                  \\
                             & b2                      & Senescence parameter                         \\
                             & b3                      & Soil water extraction rate                   \\
                             & T\textsubscript{opt}    & Optimal vegetation growth temperature        \\
                             & Phmin                   & Lower bound of optimal daily solar radiation \\
                             & Phmax                   & Upper bound of optimal daily solar radiation \\
                             & L                       & Lag (days)                                    \\
                             & h                       & fCover Transfer function parameter                                           
\end{tabular}
\end{table}

## References
Richardson, A. D., Hufkens, K., Milliman, T., Aubrecht, D. M., Chen, M., Gray, J. M., … Frolking, S. (2018). Tracking vegetation phenology across diverse North American biomes using PhenoCam imagery. Scientific Data, 5(1), 180028. https://doi.org/10.1038/sdata.2018.28  
Hufkens, K., Keenan, T. F., Flanagan, L. B., Scott, R. L., Bernacchi, C. J., Joo, E., … Richardson, A. D. (2016). Productivity of North American grasslands is increased under future climate scenarios despite rising aridity. Nature Climate Change, 6(7), 710–714. https://doi.org/10.1038/nclimate2942

### 2020-12-28 manuscript/supplement2.Rmd
add MAP scaling reference

@@
 
-The vegetaion parameter V corresponds to the fCover, which is mapped to PhenoCam G\textsubscript{cc} using the transfer function:
+The vegetaion parameter V corresponds to the fCover, which is mapped to PhenoCam G\textsubscript{cc} using the transfer function based on the relationship between annual precipitation and fractional vegetation cover (Donohue et al.
+2013).:
 
@@
 ## References
-Richardson, A.
-D., Hufkens, K., Milliman, T., Aubrecht, D.
-M., Chen, M., Gray, J.
-M., … Frolking, S.
-(2018).
-Tracking vegetation phenology across diverse North American biomes using PhenoCam imagery.
-Scientific Data, 5(1), 180028.
-https://doi.org/10.1038/sdata.2018.28
+Donohue, R.
+J., Roderick, M.
+L., McVicar, T.
+R., & Farquhar, G.
+D.
+(2013).
+Impact of CO 2 fertilization on maximum foliage cover across the globe’s warm, arid environments.
+Geophysical Research Letters, 40(12), 3031–3035.
+https://doi.org/10.1002/grl.50563
 Hufkens, K., Keenan, T.
@@
 https://doi.org/10.1038/nclimate2942
+Richardson, A.
+D., Hufkens, K., Milliman, T., Aubrecht, D.
+M., Chen, M., Gray, J.
+M., … Frolking, S.
+(2018).
+Tracking vegetation phenology across diverse North American biomes using PhenoCam imagery.
+Scientific Data, 5(1), 180028.
+https://doi.org/10.1038/sdata.2018.28

### 2020-12-28 manuscript/manuscript.Rmd
some final minor edits
- use spatial scales throughout instead of extent, iterations, etc.
- add caveat on temporal out of sample validation and wouldn't that be nice

@@
 
-A promising method is low dimensional models, which are process models with some simplified components [@choler2010; @choler2011].
+A promising method for this is low dimensional models, which are process models with some simplified components [@choler2010; @choler2011].
 For example, a low dimensional model might approximate transpiration to a function of potential evapotranspiration, soil available water, and live vegetation cover along with a single parameter.
@@
 Despite it's name the PhenoGrass model can theoretically apply to any vegetation type with a distinct growth signal in response to precipitation, as hypothesized in the original threshold-delay model [@ogle2004].
-Here we use two additional vegetation types, shrubs and agricultural plots, to test how applicable it is beyond grasslands.
+Here we evaluate two additional vegetation types, shrubs and agricultural plots, to test how applicable it is beyond grasslands.
 
@@
 We downloaded all Phenocam data with ROIs of the grasslands (GR), shrublands (SH), and agricultural (AG) vegetation types for the years 2012 to 2018, totalling 89 distinct time series and 463 site-years (Fig.
-1, Table S1).
+1, Table S2).
 As input to the PhenoGrass model we used the 3-day smoothed G\textsubscript{cc} scaled, for each ROI, from 0-1.
-In the model parameterization each ROI time series is further transformed to a fractional cover estimate using the local mean annual precipitation (MAP) combined with a scaling factor [@hufkens2016; @donohue2013].
+In the model parameterization each ROI time series is further transformed to fractional vegetation cover (\textit{fCover}, see supplmental equations) using the local mean annual precipitation (MAP) combined with a scaling factor [@hufkens2016; @donohue2013].
 
@@
 
-To find the most appropriate scale we evaluated the model using three different combinations of ecoregion and vegetation type, with 11 total model parameterizations (Fig.
+To find the most appropriate spatial scale we evaluated the model using three different combinations of ecoregion and vegetation type, with 11 total model parameterizations (Fig.
 2).
 Here we use the term “spatial scale” to refer to the combination of ecoregion/s and vegetation type/s used within each model.
-This includes using all vegetation types with an ecoregion, or all sites of a specific vegetation type from several ecoregions.
-The largest scale used all Phenocam locations described above (89 sites).
+This includes using all sites of one vegetation types within an ecoregion, or all sites of a specific vegetation type from several ecoregions (Fig.
+2).
+The largest spatial scale used all Phenocam locations described above (89 sites).
 Next were all sites, respectively, within the three vegetation types indicated by the ROI (grasslands, shrublands, and agricultural).
@@
 
-Models which exceed the threshold were subject to further evaluation.
+Models which exceeded the threshold were subject to further evaluation.
 For each model we performed a leave one out cross-validation, where the model was re-fit with one Phenocam site not included in the training data, and then evaluated against this left out site.
@@
 
-At the largest scale, where the PhenoGrass model was parameterized using all 89 sites, the model performed poorly with an \(\overline{NSE}\) value among sites of 0.31 (Table 1; Fig.
+At the largest spatial scale, where the PhenoGrass model was parameterized using all 89 sites, the model performed poorly with an \(\overline{NSE}\) value among sites of 0.31 (Table 1; Fig.
 3).
@@
 
-In all 11 iterations the PhenoGrass model tended to underestimate the highest fCover values, and to a lesser degree overpredict the lowest values (Figs.
+In all 11 models the PhenoGrass model tended to underestimate the highest fCover values, and to a lesser degree overpredict the lowest values (Figs.
 3,4).
-The best performing iterations (Grasslands in the Great Plains and Eastern Temperate forests) minimized this effect (Fig.
+The best performing models (Grasslands in the Great Plains and Eastern Temperate forests) minimized this effect (Fig.
 4).
-The worst performing iteration, Grasslands in N.A.
+The worst performing models, Grasslands in N.A.
 Deserts, had little variation in predicted fCover values, resulting in the lowest \(\overline{NSE}\) overall.
 
-The grassland vegetation type, subset to specific ecoregions, predominantly outperformed other iterations of the PhenoGrass model (Table 1).
+The grassland vegetation type, subset to specific ecoregions, predominantly outperformed models built with other spatial scales (Table 1).
 Models built using grasslands within the Eastern Temperate Forest and Great Plains ecoregions had the highest \(\overline{NSE}\) values of 0.82 and 0.69, respectively.
-Using leave one out cross-validation on these two grassland model iterations resulted in similar errors of 0.79 and 0.67 for the Eastern Temperate Forest and Great Plains, respectively.
+Using leave one out cross-validation on these two grassland models resulted in similar errors of 0.79 and 0.67 for the Eastern Temperate Forest and Great Plains, respectively.
 Though N.W.
 Forests grasslands had an in-sample \(\overline{NSE}\) just below the 0.65 threshold, the cross-validation was well below it (0.52).
-Grasslands in the North American deserts were not modelled well at any scale and had the lowest \(\overline{NSE}\) values in the entire analysis.
+Grasslands in the North American deserts were not modelled well at any spatial scale and had the lowest \(\overline{NSE}\) values in the entire analysis.
 The observed greenness patterns of these desert grasslands had extremely high variability in their magnitude and timing, with short distinct peaks in greenness and numerous off-peak fluctuations.
@@
 
-Agriculture and shrubland sites were poorly modelled at all scales.
+Agriculture and shrubland sites were poorly modelled at all spatial scales.
 Performance of agriculture within the E.
-Temperate Forest ecoregion (NSE = 0.33) improved over the All Agriculture model (NSE = 0.18), but decreased in the Great Plains (from 0.24 to 0.18).
+Temperate Forest ecoregion (\(\overline{NSE}\) = 0.33) improved over the All Agriculture model (\(\overline{NSE}\) = 0.18), but decreased in the Great Plains (\(\overline{NSE}\) from 0.24 to 0.18).
 There was only a single ecoregion with a minimum of five shrubland sites, N.A.
@@
 
-We performed an extensive evaluation of the PhenoGrass model across ecoregions and vegetation types to determine the best scale at which to parameterize and apply the model.
+We performed an extensive evaluation of the PhenoGrass model across ecoregions and vegetation types to determine the best spatial scale at which to parameterize and apply the model.
 We found the model most suitable to grassland vegetation when constrained to the ecoregion level, though it did not perform well in grasslands in the North American desert ecoregion.
-Shrublands and agriculture were not well represented by the model regardless of the scale.
+Shrublands and agriculture were not well represented by the model regardless of the spatial scale.
 Results from this study will facilitate long-term projections of grassland productivity constrained to an appropriate vegetation type and scale.
@@
 
-Shrublands were not well modelled at any scale.
+Shrublands were not well modelled at any spatial scale.
 Dryland shrubs, representing 8 of the 15 shrubland PhenoCams analysed here, coexist with grasses by accessing different pools of soil water [@weltzin2000; @muldavin2008], thus have different responses to precipitation and resulting greenness patterns [@browning2017; @yan2019].
@@
 S1), and were thus more difficult to represent in a single continental scale grassland model.
+This highlights the need for longer time series in evaluating low-dimensional models as it may take many years for a single location to experience the full range of variability.
+As the PhenoCam data archive grows, temporally out of sample validation can be done to better evaluate performance into novel conditions.
 
@@
 Here we have validated prior modelling work and highlighted its limitations.
-Newer small scale vegetation models can be validated in the same framework and applied to areas where PhenoGrass performs poorly.
+Newer vegetation models can be validated in the same framework and applied to areas where PhenoGrass performs poorly.
 This can result in a spatial ensemble where the output for any one location and vegetation type is represented by the most appropriate model.
-Our current work will allow for long-term small scale projections of grassland productivity for a large fraction of North America.
+Our current work will allow for long-term projections of grassland productivity for a large fraction of North America.

### 2020-12-28 manuscript/manuscript.Rmd
minor edits/typos

@@
 
-A promising method for this is low dimensional models, which are process models with some simplified components [@choler2010; @choler2011].
-For example, a low dimensional model might approximate transpiration to a function of potential evapotranspiration, soil available water, and live vegetation cover along with a single parameter.
-As opposed to a high dimensional model with multiple functions accounting for leaf area index, stomatal conductance, rooting depth and surface area, etc [@caylor2009; @asbjornsen2011].
-The low dimensional model is advantageous since it can generalize across broad regions with relatively few inputs.
+A promising method for this is low-dimensional models, which are process based models with some simplified components [@choler2010; @choler2011].
+For example, a low-dimensional model might approximate transpiration as a function of potential evapotranspiration, soil available water, and live vegetation cover along with a single parameter.
+As opposed to a high-dimensional model with multiple functions accounting for leaf area index, stomatal conductance, rooting depth and surface area, etc [@caylor2009; @asbjornsen2011].
+The low-dimensional model is advantageous since it can generalize across broad regions with relatively few inputs.
 Yet they are still susceptible to over-fitting to local conditions since parameters or model structure can be tied to specific locations or plant functional groups [@fisher2020].
@@
 For each model we performed a leave one out cross-validation, where the model was re-fit with one Phenocam site not included in the training data, and then evaluated against this left out site.
-In this step a scaling coefficient to link mean annual precipitation with PhenoCam G\textsubscript{cc} was held constant at the value obtained in first fitting.
+In this step a scaling coefficient to link mean annual precipitation with PhenoCam G\textsubscript{cc} was held constant at the value obtained in the first fitting.
 The resulting \(\overline{NSE}\) and \textit{F} are from all modelled sites using their respective out of sample test.
@@
 
-At the largest spatial scale, where the PhenoGrass model was parameterized using all 89 sites, the model performed poorly with an \(\overline{NSE}\) value among sites of 0.31 (Table 1; Fig.
+At the largest spatial scale, where the PhenoGrass model was parameterized using all 89 sites, the model performed poorly with an \(\overline{NSE}\) value of 0.31 (Table 1; Fig.
 3).
@@
 The observed greenness patterns of these desert grasslands had extremely high variability in their magnitude and timing, with short distinct peaks in greenness and numerous off-peak fluctuations.
-The fitted model, which minimized the mean CVME among the 5 sites, was not able to reproduce this high variability and instead produced fCover values that were severely constrained to a narrow range (Fig.
+The fitted model, which minimized \textit{F} among the 5 sites, was not able to reproduce this high variability and instead produced fCover values that were severely constrained to a narrow range (Fig.
 S1).
@@
 We performed an extensive evaluation of the PhenoGrass model across ecoregions and vegetation types to determine the best spatial scale at which to parameterize and apply the model.
-We found the model most suitable to grassland vegetation when constrained to the ecoregion level, though it did not perform well in grasslands in the North American desert ecoregion.
+We found the model most suitable to grassland vegetation when constrained to the ecoregion level, though it did not perform well in grasslands in the North American Deserts ecoregion.
 Shrublands and agriculture were not well represented by the model regardless of the spatial scale.
@@
 @hufkens2016 originally evaluated the PhenoGrass model using 14 grassland sites distributed among seven North American ecoregions.
-In their evaluation they had an average R\textsuperscript{2} of 0.71, while here the model performed poorly when using more than 1 ecoregion.
+In their evaluation they had an average R\textsuperscript{2} of 0.71, while here the model performed poorly when using more than one ecoregion.
 It’s likely that the original 14 grassland sites were ideal locations for the PhenoGrass model, since on average they have a single greenup season every year in the spring or summer (Fig.

### 2021-01-02 analysis/analyze_errors.R
Biogeosciences major revision (#13)
Submitted.

- Clarify spatial scales throughout

- R2 is now called the nash sutcliffe coeficient of efficiency, NSE,
note the math is the same

- RMSE is removed since it isn't comparable across sites. replaced
by the mean coefficient of variation of the mean absolute error,
aka the mean cvmae, aka F. This is also the loss metric used in
fitting.

- new 2nd paragraph in intro

- add text on using phenograss beyond grass

- all equations in supplement

- minor corrections throughout

@@ -8,8 +8,8 @@ all_predictions = read_csv('data/full_model_predictions.csv',
 
 site_errors = all_predictions %>%
   group_by(fitting_set, model, timeseries_id) %>%
-  summarise(rmse = sqrt(mean( (fCover_observed - fCover_predicted)^2 ,na.rm=T)),
-            r2   = 1  - sum((fCover_observed - fCover_predicted)^2,na.rm=T) / sum((fCover_observed - mean(fCover_observed,na.rm=T))^2,na.rm=T),
+  summarise(cvmae = mean(abs(fCover_observed - fCover_predicted), na.rm=T)/mean(fCover_observed, na.rm=T) ,
+            nse   = 1  - sum((fCover_observed - fCover_predicted)^2,na.rm=T) / sum((fCover_observed - mean(fCover_observed,na.rm=T))^2,na.rm=T),
             n=n(),
             percent_na=mean(is.na(fCover_observed))) %>%
   ungroup() 
@@ -17,8 +17,8 @@ site_errors = all_predictions %>%
 primary_errors = site_errors %>%
   filter(model == 'PhenoGrass') %>% # just the phenograss model in the primary things. naive errors will be in supplement
   group_by(fitting_set, model) %>%
-  summarise(rmse = round(mean(rmse),2),
-            r2 =   round(mean(r2),2),
+  summarise(mean_cvmae = round(mean(cvmae),2),
+            nse =  round(mean(nse),2),
             n_timeseries = n_distinct(timeseries_id)) %>%
   ungroup()
 

### 2021-01-02 analysis/error_figures.R

@@ -17,11 +17,11 @@ phenocam_site_info = read_csv('site_list.csv') %>%
 get_label_stats = function(df){
   # This needs to accept a group_by() output
   df %>%
-    summarise(rmse = mean(rmse),
-              r2 =   mean(r2),
+    summarise(mean_cvmae = mean(cvmae),
+              nse  =   mean(nse),
               n_timeseries = n_distinct(timeseries_id)) %>%
     ungroup() %>%
-    mutate(label_text = paste0('R2: ',round(r2,2),'\n','RMSE: ',round(rmse,2),'\n',n_timeseries,' sites'))
+    mutate(label_text = paste0('list(bar(NSE):',round(nse,2),',','italic(F):',round(mean_cvmae,2),',',n_timeseries,'~sites',')'))
 }
 
 #######################################################
@@ -39,8 +39,8 @@ allsite_veg_predictions$fitting_set = factor(allsite_veg_predictions$fitting_set
 
 allsite_veg_predictions_error_text = allsite_veg_predictions %>%
   group_by(fitting_set, timeseries_id) %>%
-  summarise(rmse = sqrt(mean( (fCover_observed - fCover_predicted)^2 ,na.rm=T)),
-            r2   = 1  - sum((fCover_observed - fCover_predicted)^2,na.rm=T) / sum((fCover_observed - mean(fCover_observed,na.rm=T))^2,na.rm=T),
+  summarise(cvmae = mean(abs(fCover_observed - fCover_predicted), na.rm=T)/mean(fCover_observed, na.rm=T),
+            nse   = 1  - sum((fCover_observed - fCover_predicted)^2,na.rm=T) / sum((fCover_observed - mean(fCover_observed,na.rm=T))^2,na.rm=T),
             n=n(),
             percent_na=mean(is.na(fCover_observed))) %>%
   ungroup() %>%
@@ -53,7 +53,7 @@ allsite_veg_error_fig = ggplot(allsite_veg_predictions, aes(x=fCover_observed, y
   geom_line(data=tibble(fCover_observed=seq(0,1,0.1), fCover_predicted=seq(0,1,0.1)),size=1, aes(color='a')) + # a 1:1 line so that a legend can be hacked in without
   geom_smooth(method='lm', se=FALSE, aes(color='b'), size=1) +                                                  # specifying a color columns
[…]

### 2021-02-19 manuscript/manuscript.Rmd
minor equation corrections
Revision round2 (#14)
* minor edits from 2nd biogeosciences revision

* latest pdf

@@
 We expand on the evaluation of the original study by using 84 PhenoCam sites, totalling 89 distinct time series, with 463 site-years of data.
-We test the model’s performance across varying combinations of North American ecoregions and vegetation types to find an optimal spatial scale in which to parameterize and apply the model.
+Using the original methods and input datasets we test the model’s performance with this expanded dataset.
+We also evaluate the model across different combinations of North American ecoregions and vegetation types to find an optimal spatial scale in which to parameterize and apply the model.
 Finally we address where the model performs poorly and how productivity projections for these areas could be implemented or improved.
@@
 Soil wilting point and field capacity were extracted at each Phenocam location from a global dataset [@soil2000].
+Despite other options being available we chose to use the same climate and soil datasets as those used in @hufkens2016 so that any differences in results can be attributed to the expanded PhenoCam dataset used here.
 
@@
 \begin{equation}
-NSE_{j} = 1- \frac{\sum_{i=1}^{n}(fCover_{obs} - fCover_{pred})^{2}}{\sum_{i=1}^{n}(fCover_{obs} - \overline{fCover_{obs}})^{2}}
+NSE_{j} = 1- \frac{\sum_{i=1}^{n}(fCover_{i,obs} - fCover_{i,pred})^{2}}{\sum_{i=1}^{n}(fCover_{i,obs} - \overline{fCover_{obs}})^{2}}
 \end{equation}
@@
 @hufkens2016 originally evaluated the PhenoGrass model using 14 grassland sites distributed among seven North American ecoregions.
-In their evaluation they had an average R\textsuperscript{2} of 0.71, while here the model performed poorly when using more than one ecoregion.
+In their evaluation they had an average NSE of 0.71, while here the model performed poorly when using more than one ecoregion.
 It’s likely that the original 14 grassland sites were ideal locations for the PhenoGrass model, since on average they have a single greenup season every year in the spring or summer (Fig.
@@
 As the PhenoCam data archive grows, temporally out of sample validation can be done to better evaluate performance into novel conditions.
+Further work could use the expanded PhenoCam timeseries to compare with flux tower derived productivity, which @hufkens2016 found was well correlated at the five original locations with available flux data.
 
